---
title: "adaptive capacity modified"
author: "Laura"
date: "11/11/2021"
output: html_document
---


##Investigating how our initial calculation of adaptive capacity compares to a broader one following the framework from Barnes ().

```{r setup, echo = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message = FALSE, error=FALSE)
```

```{r, packages}
library(tidyverse)
library(lubridate)
library(knitr)
library(grDevices)
library(extrafont)
library(ggrepel)
library(ggthemes)
library(ggplot2)
library(grid)
library(RColorBrewer)
library(ggpubr)
library(PNWColors)
library(GGally)
library(factoextra)
library(psych)
# windowsFonts(Times=windowsFont("Calibri"))
theme_sleek <- function(base_size = 12, base_family = "Calibri") {
  half_line <- base_size/2
  theme_light(base_size = 12, base_family = "Calibri") +
    theme(
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      axis.ticks.length = unit(half_line / 2.2, "pt"),
      strip.background = element_rect(fill = NA, colour = NA),
      strip.text.x = element_text(colour = "black"),
      strip.text.y = element_text(colour = "black"),
      # axis.text = element_text(colour = "grey30"),
      #axis.title = element_text(colour = "grey30"),
      #legend.title = element_text(colour = "grey30"),#, size = rel(0.9)
      legend.title = element_blank(),
      panel.border = element_rect(fill = NA),#, colour = "grey70", size = 1),
      legend.key.size = unit(0.9, "lines"),
      #legend.text = element_text(size = rel(0.7)),#, colour = "grey30"),
      legend.key = element_rect(colour = NA, fill = NA),
      legend.background = element_rect(colour = NA, fill = NA)#,
      #plot.title = element_text(colour = "grey30"),#, size = rel(1),
      #plot.subtitle = element_text(colour = "grey30")#, size = rel(.85),
     
      
    )
}
theme_set(theme_sleek())
options(scipen = 999) #turns off scientific notation
# Depends on dplyr
tickr <- function(
  data, # dataframe
  var, # column of interest
  to # break point definition 
){
  
  VAR <- enquo(var) # makes VAR a dynamic variable
  
  data %>% 
    distinct(!!VAR) %>%
    ungroup(!!VAR) %>% 
    mutate(labels = ifelse(!!VAR %in% seq(to * round(min(!!VAR) / to), max(!!VAR), to),
                           !!VAR, "")) %>%
    select(breaks = UQ(VAR), labels)
}
```

```{r data}
responses <- read.csv("../data/PNW_decoded_vulnerability.csv")
```
```{r}
#color paletes
pal2<-pnw_palette(name="Sailboat",n=2,type="discrete")
pal3<-pnw_palette(name="Sailboat",n=3,type="discrete")
pal4<-pnw_palette(name="Sailboat",n=4,type="discrete")
pal5<-pnw_palette(name="Sailboat",n=5,type="discrete")
pal6<-pnw_palette(name="Sailboat",n=6,type="discrete")
pal7<-pnw_palette(name="Sailboat",n=7,type="continuous")

```
```{r functions}
#giving exposure a scale of 0 to 1, where a fishery that will be strongly negatively affected by climate change scores a 1. Higher score means higher exposure.
exposure_calc <- function(x){
if(is.na(x)){
  return("NA")
} else if(x == "strong_neg"){
  return(1)
} else if(x == "slight_neg"){
  return(0.75)
} else if(x == "no_effect"){
  return(.5)
} else if (x == "slight_pos"){
  return(.25)
} else if (x == "strong_pos"){
  return(0)
} else if (x == "idk") {
  return("NA")
}
}

#Recoding likert answers on a scale of 0 to 1 where strongly agree equals 1. To be applied for sensitity and adaptive capacity or other applicable situations.
sensitivity_ac <- function(x){
if(is.na(x)){
  return("NA")
} else if(x == "strongly_agree"){
  return(1)
} else if(x == "somewhat_agree"){
  return(.75)
} else if(x == "neutral"){
  return(.5)
} else if (x == "somewhat_disagree"){
  return(.25)
} else if (x == "strongly_disagree"){
  return(0)
} else if (x == "I have not observed any changes") {
  return("NA")
}
}

#Recoding likert in opposite for two inverse adaptive capcity questions, strongly agree equals 1 and higher score equals higher adaptive
ac_scale <- function(x){
if(is.na(x)){
  return("NA")
} else if(x == "strongly_agree"){
  return(0)
} else if(x == "somewhat_agree"){
  return(.25)
} else if(x == "neutral"){
  return(.5)
} else if (x == "somewhat_disagree"){
  return(.75)
} else if (x == "strongly_disagree"){
  return(1)
} else if (x == "I have not observed any changes") {
  return("NA")
}
}

belief_binary <- function(x){
if(is.na(x)){
  return("NA")
} else if(x == "strongly_agree"){
  return(1)
} else if(x == "somewhat_agree"){
  return(1)
} else if(x == "neutral"){
  return(0)
} else if (x == "somewhat_disagree"){
  return(0)
} else if (x == "strongly_disagree"){
  return(0)
} else if (x == "I have not observed any changes") {
  return("NA")
}
}
```
Creating a larger adaptive capacity index that includes things overlooked in more economic driven indices like risk perception and connection with community.
```{r}
ac_barnes<-responses%>%
  select(money, vessel_owner, loan, income, other_fisheries, yrs_fishing, effective_mgmt, equitable, adapt_mgmt, connect_env, connect_comm, strong_comm, generation, believe, harm_me, harm_future, voice, const_reg, no_point, data, new_fishery, nat_resource_job, other_job) 

#add age back in when decision is made on how to score
#took vessel out
#took out number of fisheries and will just go with diversity of fisheries

ac_barnes <- ac_barnes %>% 
  mutate(money = sapply(X=ac_barnes$money, FUN=sensitivity_ac),
         effective_mgmt = sapply(X=ac_barnes$effective_mgmt, FUN=sensitivity_ac),
         equitable = sapply(X=ac_barnes$equitable, FUN=sensitivity_ac),
         connect_env = sapply(X=ac_barnes$connect_env, FUN=sensitivity_ac),
         connect_comm = sapply(X=ac_barnes$connect_comm, FUN=sensitivity_ac),
         generation = sapply(X=ac_barnes$generation, FUN=sensitivity_ac),
         believe = sapply(X=ac_barnes$believe, FUN = sensitivity_ac),
         harm_me = sapply(X=ac_barnes$harm_me, FUN = sensitivity_ac),
         harm_future = sapply(X=ac_barnes$harm_future, FUN = sensitivity_ac),
         voice = sapply(X=ac_barnes$voice, FUN=sensitivity_ac),
         no_point = sapply(X=ac_barnes$no_point, FUN=ac_scale),
         data = sapply(X=ac_barnes$data, FUN=sensitivity_ac))

#ac_barnes <- ac_barnes %>%
  #mutate(vessel = case_when(
    #vessel == "U25" ~ 0,
    #vessel == "26-35" ~ 0,
    #vessel == "36-45" ~ 0,
    #vessel == "46-55" ~ 0,
    #vessel == "56-65" ~ 1,
    #vessel == "66+" ~ 1))

ac_barnes <- ac_barnes %>%
  mutate(income = case_when(
    income == "none" ~ 0,
    income == "U10" ~ .25,
    income == "10-25" ~ .5,
    income == "25-50" ~ .75,
    income == "50+" ~ 1))

ac_barnes <- ac_barnes %>%
  mutate(other_fisheries = case_when(
    other_fisheries == "yes" ~ 1,
    other_fisheries == "no" ~ 0))

#fisheries that generate greater than 25% of income in next section

#ac_barnes <- ac_barnes %>%
  #mutate(num_fisheries = case_when(
    #num_fisheries == "1" ~ 0,
    #num_fisheries == "2" ~ 1,
    #num_fisheries == "3" ~ 1,
    #num_fisheries == "4" ~ 1,
    #num_fisheries == "4+" ~ 1
    #))

ac_barnes <- ac_barnes %>%
  mutate(yrs_fishing = case_when(
    yrs_fishing == "0-5" ~ .25,
    yrs_fishing == "5-15" ~ .5,
    yrs_fishing == "15-25" ~ .75,
    yrs_fishing == "25+" ~ 1))
```

Using the responses to how many fisheries generate more than 25% of your income to get a measure of income eveness (Holland and Kasperson 2016?).
```{r}
diverse <- responses %>%
  select(CPS_income, dungeness_crab_income, groundfish_income, hake_income, HMS_income, salmon_income, scallops_income, urchin_income, shrimp_income, squid_income, other_income)

diverse$fisheries_diversity = rowSums(diverse)

diverse <- diverse %>%
  mutate(fisheries_diversity = case_when(
    fisheries_diversity == "1" ~ .25,
    fisheries_diversity == "2" ~ .5,
    fisheries_diversity == "3" ~ .75,
    fisheries_diversity == "4" ~ 1))

ac_barnes$fishery_diversity = diverse$fisheries_diversity
```
Making one climate variable, score is higher if you answered yes to all questions - believe, harm me, harm future, and then taking those columns out of the index
```{r}

ac_barnes<- ac_barnes %>% 
  mutate(climate = rowMeans(ac_barnes[,14:16]))

ac_barnes<- ac_barnes %>% select(-c(believe, harm_me, harm_future))

```
Add a number of gear types for flexibility
```{r}
gears <- responses %>%
  select(longline, midwater_trwl, bottom_trwl, pots, purse_seine, gillnet, troll, other_gear)

gears$geartotal = rowSums(gears)

gears <- gears %>%
  mutate(geartotal = case_when(
    geartotal == "1" ~ .25,
    geartotal == "2" ~ .5,
    geartotal == "3" ~ .75,
    geartotal == "4" ~ 1))

ac_barnes$gear = gears$geartotal
```
Add worry level for relative risk percetion of climate vs other issues
```{r}

#need to run some chunks in the concerns code to get this

worryrank<-worries %>%
  select(marine_env, fishing_ops, community_infras, personal)

worryrank <- worryrank %>%
  mutate(climaterank = case_when(
  marine_env > fishing_ops &  marine_env > community_infras &  marine_env > personal ~ 1,
  marine_env < fishing_ops &  marine_env < community_infras &  marine_env < personal ~ 0))

worryrank<-worryrank%>%
  replace_na(list(climaterank = .5))

ac_barnes$worryrank = worryrank$climaterank
```

Are the variables in this new index correlated with each other?
```{r}
indicator_corr<-ggcorr(ac_barnes[,1:24])

indicator_corr

indicator_cor<-cor(ac_barnes[,1:24], method = "pearson")

cor.plot(ac_barnes, numbers = TRUE)
```
Now taking the average of everything to get a new value for adaptive capacity. Range is still 0 to 1. 
```{r new ac index}
ac_barnes$new_ac = rowMeans(ac_barnes)
```

Putting all the indicators into the domains
```{r}
ac_barnes <- ac_barnes %>%
  rowwise()%>%
  mutate(assets = mean(c(money,vessel_owner, loan)),
         flexibility = mean(c(other_fisheries, income , gear, new_fishery, other_job)), #fishery_diversity, 
         organization = mean(c(effective_mgmt, equitable, adapt_mgmt, connect_comm, strong_comm)),
         learning = mean(c(generation, data)),
         sociocog = mean(c(climate, worryrank)),
         agency = mean(c(voice, const_reg, no_point)))

domainsmod <- ac_barnes %>%
  select(assets, flexibility, organization, learning, sociocog, agency)

domainsmod$ac = rowSums(domainsmod)
```
PCA of ac by domains modified
```{r}

ac_dom_pca2<-prcomp(domainsmod[,1:6], scale. = FALSE)
summary(ac_dom_pca2)

ac_dom_pca2$rotation
ac_dom_pca2$x
str(ac_dom_pca2)

#eigenvalues
eig_new2<-get_eig(ac_dom_pca2)
eig_new2

fviz_eig(ac_dom_pca2)

#results for variable
var_new2 <- get_pca_var(ac_dom_pca2)
var_new2$coord          # Coordinates
var_new2$contrib        # Contributions to the PCs
var_new2$cos2           # Quality of representation 

dom_pca_plot2<-fviz_pca_var(ac_dom_pca2,
             col.var = "contrib", # Color by contributions to the PC
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE     # Avoid text overlapping
             )
dom_pca_plot2

ggsave(plot = dom_pca_plot2, file = paste0("../figures/ac/dom_pca_biplot2.png"))

domainsmod$fisheries = responses$listoffisheries
groups <- as.factor(responses$quadrant)

quadgroups<-fviz_pca_ind(ac_dom_pca2,
             col.ind = groups,
             #palette = pnw_palette("Bay", 62, type = "continuous"),
             addEllipses = TRUE, # Concentration ellipses
             ellipse.type = "confidence",
             legend.title = "Groups",
             repel = TRUE
             )

ggsave(plot = quadgroups, file = paste0("../figures/ac/quadgroups.png"))

fviz_pca_var(ac_dom_pca2, col.var = "black")
library("corrplot")
corrplot(var_new2$cos2, is.corr = FALSE)
```
Flexibility isn't well represented with these pc's, how does it correlate with some of the self-reported measures of flexibility from the survey?
```{r}
flexindex<-responses %>%
  select(new_fishery, nat_resource_job, other_job)

flexindex$avgflex = rowMeans(flexindex)

cor.test(domainsmod$flexibility, flexindex$avgflex, method = "pearson")

flex<-data.frame(domainsmod$flexibility, flexindex$avgflex)

ggscatter(flex, x = "domainsmod.flexibility", y = "flexindex.avgflex", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "domain", ylab = "survey question")

flexincluded<-ac_barnes %>%
  select(income, other_fisheries, fishery_diversity, gear)
flexincluded$new_fishery = responses$new_fishery
flexincluded$natresource = responses$nat_resource_job
flexincluded$other_job = responses$other_job
flexincluded$flex = domainsmod$flexibility

corr.test(flexincluded)
```
```{r}
groupsage <- as.factor(responses$age)

agegroups<-fviz_pca_ind(ac_dom_pca,
             col.ind = groupsage,
             #palette = pnw_palette("Bay", 62, type = "continuous"),
             addEllipses = TRUE, # Concentration ellipses
             ellipse.type = "confidence",
             legend.title = "Groups"
             )

agegroups

groupsves <- as.factor(responses$vessel)

vesgroups<-fviz_pca_ind(ac_dom_pca,
             col.ind = groupsves,
             #palette = pnw_palette("Bay", 62, type = "continuous"),
             addEllipses = TRUE, # Concentration ellipses
             ellipse.type = "confidence",
             legend.title = "Groups"
             )

vesgroups

groupsab <- as.factor(responses$ability)

abilgroups<-fviz_pca_ind(ac_dom_pca,
             col.ind = groupsab,
             #palette = pnw_palette("Bay", 62, type = "continuous"),
             addEllipses = TRUE, # Concentration ellipses
             ellipse.type = "confidence",
             legend.title = "Groups"
             )

abilgroups

```

#trying the same plots with ggbiplot to see if the elipses turn up better
```{r}
#library(ggbiplot)
#ggbiplot(ac_dom_pca, ellipse = TRUE, groups = groupsab)
#this one is potentially interesting
```

#trying out the psych package
```{r}
pcp3<-principal(domainsmod[,1:6], 3, n.obs = 162, rotate = "Varimax")
pcp3

fc3<-fa(ac_barnes[,1:18], 3, n.obs = 162, fm = "wls")
print(fc3, cut = 0, digits = 2)
plot(fc3)
fa.diagram(fc3)

bassAckward(ac_barnes[,1:18])

fa.parallel(ac_barnes[,1:18])
vss(ac_barnes[,1:18])


fc6<-fa(ac_barnes[,1:18], 6, n.obs = 162, fm = "wls")
print(fc6, cut = 0, digits = 2)
plot(fc6)
fa.diagram(fc6)

om<-omega(ac_barnes[,1:18], n.obs = 162)
```

```{r}
domainsmod$age = responses$age
domainsmod$vessellen = responses$vessel
domainsmod$experience = responses$yrs_fishing
domainsmod$fisheries = responses$listoffisheries

pairs.panels(domainsmod [,1:7], pch = '.')

pairs(domainsmod[,1:7], smooth = TRUE, scale = FALSE, density=TRUE,ellipses=TRUE,
digits = 2,method="pearson", pch = 20, lm=FALSE,cor=TRUE,jiggle=FALSE,factor=2,
hist.col="cyan",show.points=TRUE,rug=TRUE, breaks = "Sturges",cex.cor=1,wt=NULL,
smoother=FALSE,stars=FALSE,ci=FALSE,alpha=.05)
```

##anova for adaptive capacity by age
```{r echo=FALSE, warning=FALSE}
#can copy and paste to look at age and vulnerability or any number of factors
#ac/vulnerability and amount of income gained outside of fishing
#ac/vulnerability and role in the industry, etc


age_ac<-ggboxplot(domainsmod, x = "age", y = "ac",
          color = "age", palette = c("#00AFBB", "#E7B800", "#FC4E07", "#00AFBB", "#E7B800", "#FC4E07"),
          order = c("U30", "30-40", "40-50", "50-60", "60-70", "70+"),
          ylab = "adaptive capacity", xlab = "age")

ggsave(plot = age_ac, file = paste0("../figures/ac/age_ac.png"))
age_ac

#consolodate age groups
domainsmod <- domainsmod %>%
  mutate(age2 = case_when(
    age == "U30" ~ "U60",
    age == "30-40" ~ "U60",
    age == "40-50" ~ "U60",
    age == "50-60" ~ "U60",
    age == "60-70" ~ "over60",
    age == "70+" ~ "over60",
  ))

t.test(ac ~ age2, data = domainsmod)

#ac
#anova
age_aov<-aov(ac ~ age, data = domainsmod)
summary(age_aov)

#tukey HSD of anova
age_tukey<-TukeyHSD(age_aov)
age_tukey
```

##anova for adaptive capacity by vessel length
```{r echo=FALSE, warning=FALSE}
#can copy and paste to look at age and vulnerability or any number of factors
#ac/vulnerability and amount of income gained outside of fishing
#ac/vulnerability and role in the industry, etc


vl_ac<-ggboxplot(domainsmod, x = "vessellen", y = "ac",
          color = "vessellen", palette = c("#00AFBB", "#E7B800", "#FC4E07", "#00AFBB", "#E7B800", "#FC4E07"),
          order = c("U25", "26-35", "36-45", "46-55", "56-65", "66+"),
          ylab = "adaptive capacity", xlab = "vessel length")

#ggsave(plot = age_ac, file = paste0("../figures/ac/age_ac.png"))
vl_ac

#ac
#anova
vl_aov<-aov(ac ~ vessellen, data = domainsmod)
summary(vl_aov)

#tukey HSD of anova
vl_tukey<-TukeyHSD(vl_aov)
vl_tukey

#consolodate vessel size groups
domainsmod <- domainsmod %>%
  mutate(vessel2 = case_when(
    vessellen == "U25" ~ "65 and under",
    vessellen == "26-35" ~ "65 and under",
    vessellen == "36-45" ~ "65 and under",
    vessellen == "46-55" ~ "65 and under",
    vessellen == "56-65" ~ "65 and under",
    vessellen == "66+" ~ "66 and over",
  ))

t.test(ac ~ vessel2, data = domainsmod)

vesselgroup_ac<-ggboxplot(domainsmod, x = "vessel2", y = "ac",
          fill = "vessel2", palette = pal2,
          order = c("65 and under", "66 and over"),
          ylab = "adaptive capacity", xlab = "vessel size") +
          theme(legend.position = "none")

ggsave(plot = vesselgroup_ac, file = paste0("../figures/ac/vesselsgrouped_ac.png"))
vesselgroup_ac
```

AC by state?
```{r}
domainsmod$portstate = responses$port_state_code
domainsmod$homestate = responses$state_code
#anova port state
ps_aov<-aov(ac ~ portstate, data = domainsmod)
summary(ps_aov)

#anova home state
hs_aov<-aov(ac ~ homestate, data = domainsmod)
summary(hs_aov)

#tukey HSD of anova
#vl_tukey<-TukeyHSD(vl_aov)
#vl_tukey
```

Looking at ac geographically as that has been a factor in range shifts. Load zip code package and get lat and long.
```{r zipcode data}
#download zipcode data to match states to responses data by zip code (could add county or other identifying information from these data)
#for survey versions in the US

ZipCodeSourceFile = "http://download.geonames.org/export/zip/US.zip"
temp <- tempfile()
download.file(ZipCodeSourceFile , temp)
ZipCodes <- read.table(unz(temp, "US.txt"), sep="\t")
unlink(temp)
names(ZipCodes) = c("CountryCode", "home_zip", "PlaceName", 
                    "state", "state_code", "County", "AdminCode2", 
                    "AdminName3", "AdminCode3", "latitude", "longitude", "accuracy")

portzips<-ZipCodes %>% 
  select(home_zip, County, latitude, longitude) %>% 
  mutate(home_zip = as.integer(home_zip))

names(portzips) = c("port_zip", "county", "latitude", "longitude")

responses <- responses %>% 
  left_join(portzips)

```

```{r}

domainsmod$county = responses$county

domainsmod$county <- as.factor(domainsmod$county)

#library(dplyr)
countyreport<-domainsmod%>%
  group_by(county) %>%
  dplyr:: summarise(avg_ac = mean(ac),
            count = n())

domainsmod$portlat<-responses$latitude
domainsmod$portlong<-responses$longitude

domainsmod<-domainsmod %>%
  mutate(region = case_when(portlat < 38.1 ~ "south_pr", #south of point reyes
                            portlat >= 38.1 & portlat < 44.7 ~ "pr_newport", #point reyes to newport, or
                            portlat >= 44.7 & portlat < 46.2 ~ "newport_astoria", #newport to astoria
                            portlat >= 46.2 & portlong < -123.38 ~ "wacoast", #north of astoria and west of and including port angeles
                            #portlat >= 47.05 & portlong < -123.38 ~ "nowa", #north of ocean shores and west of and including port angeles
                            portlat >= 46.2 & portlong >= -123.38 ~ "esjf_ps"))
#east of port angeles and into puget sound

#filter out the outlier
#domains<-domains %>%
  #filter(ac > 1.2)

region_aov<-aov(ac ~ region, data = domainsmod)
summary(region_aov)

regionscatter <- ggplot(domainsmod, aes(x = portlat, y = ac, color = region))+
  geom_point()

regionscatter

domainsmod$salmon = responses$salmon
domainsmod$salmon <- as.factor(domainsmod$salmon)

salmon_scatter <- ggplot(domainsmod, aes(x = portlat, y = ac, color = salmon))+
  geom_point()
salmon_scatter

domainsmod$diversity<-as.factor(domainsmod$diversity)
diversity_scatter <- ggplot(domainsmod, aes(x = portlat, y = ac, color = diversity))+
  geom_point()
diversity_scatter

ggplot(domainsmod, aes(x = diversity, y = ac, color = region))+
  geom_point()

#regionrisk <- ggplot(domains, aes(x = portlat, y = risk, color = region)) +
  #geom_point() +
  #geom_smooth(method = lm)
#regionrisk
                  
```
continuing to investigate the region stuff
```{r}

#assumptions of anova
library(car)
leveneTest(ac ~ region, data = domainsmod)

aov.residuals<-residuals(object = region_aov)
shapiro.test(x=aov.residuals)

regionsmod<-domainsmod %>%
  group_by(region) %>%
  dplyr:: summarise(avg_ac = mean(ac),
                    count = n())

#anova
region_aov<-aov(ac ~ region, data = domainsmod)
summary(region_aov)

#tukey HSD of anova
region_tukey<-TukeyHSD(region_aov)
region_tukey

region_ac<-ggboxplot(domainsmod, x = "region", y = "ac",
          fill = "region", palette = pal5,
          order = c("south_pr", "pr_newport", "newport_astoria", "wacoast", "esjf_ps"),
          ylab = "adaptive capacity", xlab = "region")+
  theme(legend.position = "none")
  #stat.summary(fun.y = mean, geom="point", shape = 20, size = 4, color = "black", fill="black")

region_ac

ggsave(plot = region_ac, file = paste0("../figures/ac/_ac.png"))
```

Putting regions on the pca graph
```{r}
rcomp<-domainsmod %>%
  select(assets, flexibility, organization, learning, sociocog, agency, region, ac)

rsum<-rcomp %>%
  group_by(region) %>%
  summarise(assets = mean(assets),
            #assetssd = sd(assets),
            flexibility = mean(flexibility),
            #flexibilitysd = sd(flexibility),
            organization = mean(organization),
            #organizationsd = sd(organization),
            learning = mean(learning),
            #learningsd = sd(learning),
            sociocog = mean(sociocog),
            #sociocogsd = sd(sociocog),
            agency = mean(agency),
            ac = mean(ac))
            #agencysd = sd(agency))
           

write.csv(rsum, "../figures/ac/regionsummary.csv")

table(domainsmod$region)


#library(reshape2)

rsumlong<-melt(rsum, id = "region")
rsumlong$region<-as.factor(rsumlong$region)


regionbar<-ggplot(rsumlong, aes(x=variable, y=value, fill = region)) +
  geom_bar(position = "dodge", stat = "identity")+
  scale_fill_manual(values = pal5)
  #geom_errorbar( aes(x=variable, ymin=value-sd, ymax=value+sd), width=0.4, colour="black", alpha=0.9, size=1.3)
  
  ggsave(plot = regionbar, file = paste0("../figures/ac/regioncompoment_ac.png"))

ggplot(rsumlong, aes(x=variable, y=value, fill = region)) +
  geom_violin(position = "dodge", alpha=0.5, outlier.colour="transparent")+
  scale_fill_manual(values = pal5)


bb<-predict(ac_dom_pca2, newdata = rsum)

rownames(bb)<-rsum$region

fviz_add(dom_pca_plot2, bb, color = "navy", repel = TRUE)
```
shit ton of anovas but domain of ac by region
```{r}
ass_aov<-aov(assets ~ region, data = domainsmod)
summary(ass_aov)

flex_aov<-aov(flexibility ~ region, data = domainsmod)
summary(flex_aov)
org_aov<-aov(organization ~ region, data = domainsmod)
summary(org_aov)
learn_aov<-aov(learning ~ region, data = domainsmod)
summary(learn_aov)

soc_aov<-aov(sociocog ~ region, data = domainsmod)
summary(soc_aov)

TukeyHSD(soc_aov)

agency_aov<-aov(agency ~ region, data = domainsmod)
summary(agency_aov)

TukeyHSD(agency_aov)

#asumptions
leveneTest(sociocog ~ region, data = domainsmod)

aov.residuals<-residuals(object = soc_aov)
shapiro.test(x=aov.residuals)

#sociocog meets levene test but fails shapiro, use kruskal wallis
kruskal.test(sociocog ~ region, data = domainsmod)

#agency assumptions
leveneTest(agency ~ region, data = domainsmod)
aov.residuals<-residuals(object = agency_aov)
shapiro.test(x=aov.residuals)

#agency also fails run kruskal
kruskal.test(agency ~ region, data = domainsmod)
```

Consoldating fishery categories for the few random things
```{r}
domainsmod<- domainsmod %>%
  mutate(fishery = case_when(
    fisheries == "dungeness crab, salmon, salmon enhancement" ~ "dungeness crab, salmon",
    fisheries == "dungeness crab, Purchasing of seafood product" ~ "dungeness crab"
  ))
```



Seperate all the fisheries in their own column
```{r}
sep_fish <- domainsmod %>% separate(fisheries, c("f1", "f2", "f3", "f4", "f5", "f6"), ",", extra = "drop", fill = "right")

sep_fish <- sep_fish %>%
  select(f1, f2, f3, f4, f5, f6)

sep_fish <- sep_fish %>% add_column(ID = responses$X, .before = "f1")

library(reshape2)
fisherylist<- melt(sep_fish, id.vars = "ID")

fisherylist <- arrange(fisherylist, ID) 

write.csv(fisherylist, paste0("../data/fisherylist_forMary.csv"))

```
adding on species risk
```{r}
fisherylist$value = trimws(fisherylist$value)

fisherylist <- fisherylist %>%
  mutate(climaterisk = case_when(
    value == "salmon" ~ 0.3796,
    value == "CPS" ~ 0.2927,
    value == "dungeness crab" ~ 0.2341,
    value == "sea urchin" ~ 0.3455,
    value == "sea cucumber" ~ 0.2885,
    value == "HMS" ~ 0.3477,
    value == "hake" ~ 0.4163,
    value == "groundfish" ~ 0.2646,
    value == "shrimp" ~ 0.2656,
    value == "squid" ~ 0.2575,
    value == "geoduck" ~ 0.2689,
    value == "razor clams" ~ 0.2806,
    value == "halibut" ~ 0.2985,
    value == "California halibut" ~ 0.2985,
    value == "smelt" ~ .4100,
    value == "black cod" ~ .3834,
    value == "sablefish" ~ .3834,
    value == "albacore" ~ .3522,
    value == "white sea bass" ~ 0.2982,
    value == "crawfish" ~ 0.2))

indv_risk<-fisherylist %>%
  group_by(ID) %>%
  dplyr::summarise(avg_risk = mean(climaterisk, na.rm = TRUE))

domainsmod$risk = indv_risk$avg_risk
```
looking at risk as calculated by koehn and where you fall out on the ac pca plot
```{r}
#ggplot(indv_risk, aes(x=avg_risk)) +
         #geom_histogram(binwidth = .03)

indv_risk <- indv_risk %>%
  mutate(rank = percent_rank(avg_risk))

indv_risk$quartile = indv_risk$rank

indv_risk <- indv_risk %>%
  mutate(quartile = case_when(
    quartile >= .75 ~ "highest risk",
    quartile >= .5 ~ "second risk",
    quartile >= .25 ~ "third risk",
    quartile >= 0 ~ "lowest risk"
  ))

groupsrisk <- as.factor(indv_risk$quartile)

riskgroups<-fviz_pca_ind(ac_dom_pca2,
             col.ind = groupsrisk,
             #palette = pnw_palette("Bay", 62, type = "continuous"),
             addEllipses = TRUE, # Concentration ellipses
             ellipse.type = "confidence",
             legend.title = "Groups"
             )

riskgroups
```
Average domain ac by fishery, using old way for now
```{r}
fisheryac = matrix(NA, nrow = 11, ncol = 8)

colnames(fisheryac) <- c("fishery", "assets", "flexibility", "organization", "learning", "sociocog", "agency", "ac")

fisheryac[,1] = c("CPS", "dungeness crab", "geoduck", "groundfish", "HMS", "razor clam", "salmon", "sea cucumber", "sea urchin", "shrimp", "squid")

#subset to the fishery and avg vulnerability by people who fish for that fishery
for(i in 1:nrow(fisheryac)) {
  index = which(grepl(fisheryac[i,1], domainsmod$fisheries) == TRUE) 
  #which contain the fishery specified at i, 1 in the matrix
  #so for i = 1, this gives you which rows (which fishermen) fish for CPS
  fisheryac[i,2] = mean(domainsmod$assets[index], na.rm = TRUE)
  fisheryac[i,3] = mean(domainsmod$flexibility[index], na.rm = TRUE)
  fisheryac[i,4] = mean(domainsmod$organization[index], na.rm = TRUE)
  fisheryac[i,5] = mean(domainsmod$learning[index], na.rm = TRUE)
  fisheryac[i,6] = mean(domainsmod$sociocog[index], na.rm = TRUE)
  fisheryac[i,7] = mean(domainsmod$agency[index], na.rm = TRUE)
  fisheryac[i,8] = mean(domainsmod$ac[index], na.rm = TRUE)
}

#change to a dataframe and to numeric
fisheryac_summary<-as.data.frame(fisheryac)
fisheryac_summary[,2:8]<-sapply(fisheryac_summary[,2:8], as.numeric)
#fishery_summary<-round(fishery_summary[,2:13], digits = 2)

fisheryac_summary
```
pca by fishery
```{r}
fish_pca<-prcomp(fisheryac_summary[,2:7], scale. = FALSE)
summary(fish_pca)

fish_pca$rotation
fish_pca$x
str(fish_pca)

#eigenvalues
eig_new<-get_eig(fish_pca)
eig_new

fviz_eig(fish_pca)

#results for variable
var_new <- get_pca_var(fish_pca)
var_new$coord          # Coordinates
var_new$contrib        # Contributions to the PCs
var_new$cos2           # Quality of representation 

fish_pca_plot<-fviz_pca_var(fish_pca,
             col.var = "contrib", # Color by contributions to the PC
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE     # Avoid text overlapping
             )
fish_pca_plot

ggsave(plot = fish_pca_plot, file = paste0("../figures/ac/fish_pca_biplot.png"))

fviz_pca_biplot(fish_pca, repel = TRUE,
                col.var = "#2E9FDF", # Variables color
                col.ind = "#696969"  # Individuals color
                )
```
#new code to try pca groups
```{r}
clusterbiplot<-fviz_pca_biplot(ac_dom_pca2,
             geom.ind = "point", # show points only (nbut not "text")
             col.ind = domainsmod$groups, # color by groups
             palette = pal5,
             col.var = "black", repel = TRUE,
             addEllipses = TRUE, # Concentration ellipses
             legend.title = "Groups"
             )

ggsave(plot = clusterbiplot, file = paste0("../figures/ac/cluster_pca_biplot.png"))


#ggplot(domainsmod, aes(x=Commercial.Engagement, y = ac))+
  #geom_point()+
  #geom_smooth()
```

```{r}

domains$ID = responses$X

fisheryac_summary <-rename(fisheryac_summary, ID = fishery)

domainsplus<- domains %>%
  select(ID, assets, flexibility, organization, learning, sociocog, agency, ac)

domainsplus <- rbind(domainsplus, fisheryac_summary)

```
trying to put the fisheries on the pca graph
```{r}
rownames(fisheryac_summary)<-fisheryac_summary$fishery
abc<- predict(ac_dom_pca, newdata = fisheryac_summary[,2:7])
abc

fviz_add(dom_pca_plot, abc, color = "navy", repel = TRUE)


```

```{r}

domains$fisheries25 = responses$fisheries25

acfish25 = domains %>%
  group_by(fisheries25) %>%
  dplyr::summarise(assets = mean(assets, na.rm = TRUE),
                   flexibility = mean(flexibility, na.rm = TRUE),
                   learning = mean(learning),
                   organization = mean(organization),
                   socio = mean(sociocog, na.rm = TRUE),
                   agency = mean(agency),
                   ac = mean(ac),
                   count = n()
)

acfish25

acfish<-t(acfish25)

acfish<-as.data.frame(acfish)

colnames(acfish)<-acfish[1,]
row.names(acfish25)<-acfish25$fisheries25

bb<-acfish25 %>%
  select(assets, flexibility, learning, organization, socio, agency, ac)

rownames(bb)<-acfish25$fisheries25

dist<-(dist(bb))

hc<-hclust(dist)
plot(hc)
```
```{r}
#who is on the big boats?

bigboat<-responses %>%
  filter(vessel == "66+")

seattle<-responses %>%
  filter(homeport == "Seattle")
```
Do poeple with better diveristy think they have better adaptive capcity?
```{r}

domainsmod$diversity = rowSums(diverse[,1:11])
domainsmod$diversity <- as.factor(domainsmod$diversity)
domainsmod$numfish <-responses$num_fisheries

corr.test(domainsmod[,c(2,29)])
pairs.panels(domainsmod[,c(1:7,22, 23, 31)])

domainsmod$Commercial.Engagement<-as.numeric(domainsmod$Commercial.Engagement)
domainsmod$Commercial.Reliance<-as.numeric(domainsmod$Commercial.Reliance)
domainsmod$risk<-indv_risk$avg_risk

domainsmod <- domainsmod %>%
  mutate(diversity = case_when(
    diversity == "1" ~ "1",
    diversity =="2" ~"2",
    diversity == "3" ~ "3 or 4",
    diversity == "4" ~ "3 or 4"
  ))

diverse_ac<-ggboxplot(domainsmod, x = "diversity", y = "ac",
          color = "diversity", palette = c("#00AFBB", "#E7B800", "#FC4E07"),
          #order = c("U30", "30-40", "40-50", "50-60", "60-70", "70+"),
          ylab = "adaptive capacity", xlab = "age")

diverse_ac

#anova
diversity_aov<-aov(ac ~ diversity, data = domainsmod)
summary(diversity_aov)

#tukey HSD of anova
diversity_tukey<-TukeyHSD(diversity_aov)
diversity_tukey
```

Looking at regional diversity
```{r}
rd<-responses %>%
  select(socal, cencal, norcal, or_coast, cr, wa_coast, pugetsound_SJF)

rd$total_areas <- rowSums(rd)

domainsmod$total_areas = rd$total_areas

ggplot(domainsmod, aes(x = total_areas, y = ac, color = region))+
  geom_point()

areas_aov<-aov(ac ~ total_areas, data = domainsmod)
summary(areas_aov)

rd$ac = domainsmod$ac
rd<-rd %>%
  mutate(area_cat = case_when(
    total_areas == 1 ~ 1,
    total_areas == 2 ~ 2,
    total_areas == 3 ~ 3,
    total_areas == 4 ~ 4,
    total_areas == 5 ~ 4,
    total_areas == 6 ~ 4
  ))

rd<-rd %>%
  mutate(area_binary = case_when(
    total_areas == 1 ~ 1,
    total_areas == 2 ~ 2,
    total_areas == 3 ~ 2,
    total_areas == 4 ~ 2,
    total_areas == 5 ~ 2,
    total_areas == 6 ~ 2
  ))

rd$area_binary<-as.factor(rd$area_binary)
aaov<-aov(ac ~ area_binary, data = rd)
summary(aaov)
```

```{r}
cor.plot(domainsmod[])
```

#STOP
#write relevant things above here!!
#Below here are things that might be useful but are not currently being used
Comparing the first set of ac numbers with the new ones.
```{r}
ac_barnes$old_ac = responses$indv_ac

ac <- ac_barnes %>%
  select(old_ac, new_ac)

cor.test(ac_barnes$old_ac, ac_barnes$new_ac, method = "pearson")
cor.test(ac_barnes$old_ac, ac_barnes$new_ac, method = "spearman")

accor<-ggscatter(ac_barnes, x = "old_ac", y = "new_ac", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "original ac index", ylab = "new ac index")

ggsave(plot = accor, file = paste0("../figures/ac/ac_index_cor.png"))

corplot<-ggpairs(ac, title = "correlation of different ac indices")

corplot

```

#likert example if needed
```{r}
#summary of all answers to begin with

acl<- ac_barnes %>%
  select(money, )

#add statements as column headers
q37_statements<-c(
  "I believe climate change is occurring",
  "Climate change will harm me personally",
  "Climate change will harm future generations",
  "If I had a choice I would leave fishing",
  "It is a big risk to move into a new fishery",
  "There is no point in preparing for climate change since we don't know exactly what is going to happen",
  "There will not be enough fish to continue to operate in my main fishery in 20 years")
colnames(q37)<-q37_statements

#setting the levels
#levels need to be set since each possible answer does not appear in each column
mylevels <- c('strongly_disagree', 'somewhat_disagree', 'neutral', 'somewhat_agree', 'strongly_agree')
for(i in seq_along(q37)) {
  q37[,i] <- factor(q37[,i], levels=mylevels)
}

future_thoughts<-likert(q37)
summary(future_thoughts)
climate_perspectives<-plot(future_thoughts, colors=rev(pnw_palette("Sailboat", 5)))

ggsave(plot = climate_perspectives, height = 7, width = 10, file = paste0("../figures/wellbeing/climate_perspectives.png"))

#by state
future_thoughts_states<-likert(q37, grouping = responses$state_code)
climate_perspectives_states_plot<-plot(future_thoughts_states, colors=c('orange','darkorange', "gray", 'darkblue','blue'))

ggsave(plot = climate_perspectives_states_plot, height = 12, width = 10, file = paste0("../figures/wellbeing/climate_perspectives_state.png"))

```

Variables driving old index
```{r}
orig_ac<-responses %>%
  select(new_fishery,nat_resource_job, other_job, loan, distance, strong_comm, climate_refugees, adapt_mgmt, const_reg)

orig_ac_pca <- prcomp(orig_ac, scale. = FALSE)
summary(orig_ac_pca)

orig_ac_pca$rotation
orig_ac_pca$x
str(orig_ac_pca)

#eigenvalues
eig_orig<-get_eig(orig_ac_pca)
eig_orig

fviz_eig(orig_ac_pca)

#results for variable
var_orig <- get_pca_var(orig_ac_pca)
var_orig$coord          # Coordinates
var_orig$contrib        # Contributions to the PCs
var_orig$cos2           # Quality of representation 

orig_pca_plot<-fviz_pca_var(orig_ac_pca,
             col.var = "contrib", # Color by contributions to the PC
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE     # Avoid text overlapping
             )

orig_pca_plot

ggsave(plot = orig_pca_plot, file = paste0("../figures/ac/orig_index_pca_biplot.png"))
```
Average new and orig ac by fishery
```{r}
fisheryac = matrix(NA, nrow = 13, ncol = 3)
colnames(fisheryac) <- c("fishery", "orig_ac", "new_ac")
fisheryac[,1] = c("CPS", "dungeness crab", "geoduck", "groundfish", "hake", "HMS", "razor clam", "salmon", "scallops", "sea cucumber", "sea urchin", "shrimp", "squid")

#subset to the fishery and avg vulnerability by people who fish for that fishery
for(i in 1:nrow(fisheryac)) {
  index = which(grepl(fisheryac[i,1], responses$listoffisheries) == TRUE) 
  #which contain the fishery specified at i, 1 in the matrix
  #so for i = 1, this gives you which rows (which fishermen) fish for CPS
  fisheryac[i,2] = mean(ac_barnes$old_ac[index], na.rm = TRUE)
  fisheryac[i,3] = mean(ac_barnes$new_ac[index], na.rm = TRUE)
}

#change to a dataframe and to numeric
fisheryac_summary<-as.data.frame(fisheryac)
fisheryac_summary[,2:3]<-sapply(fisheryac_summary[,2:3], as.numeric)
#fishery_summary<-round(fishery_summary[,2:13], digits = 2)

fisheryac_summary

#visualize difference using lollipop chart
library(hrbrthemes)
fd<-ggplot(fisheryac_summary) +
  geom_segment( aes(x=fishery, xend=fishery, y=orig_ac, yend=new_ac), color="black") +
  geom_point( aes(x=fishery, y=orig_ac), color=rgb(0.2,0.7,0.1,0.5), size=3 ) +
  geom_point( aes(x=fishery, y=new_ac), color=rgb(0.7,0.2,0.1,0.5), size=3 ) +
  coord_flip()+
  theme_ipsum() +
  theme(
    legend.position = "none",
  ) +
  xlab("") +
  ylab("average adaptive capacity")

ggsave(plot = fd, file = paste0("../figures/ac/fishery_difference.png"))

fisheryac_summary <- fisheryac_summary %>%
  add_column(pct_change = NA)

fisheryac_summary <- fisheryac_summary %>%
  mutate(pct_change = ((new_ac - orig_ac)/orig_ac)*100)

pct_diff<-ggplot(fisheryac_summary, aes(x=fishery, y=pct_change)) +
  geom_bar(stat = "identity", color = "navy")

ggsave(plot = pct_diff, file = paste0("../figures/ac/fishery_difference_pct.png"))
```

##anova for adaptive capacity by age
```{r echo=FALSE, warning=FALSE}
#can copy and paste to look at age and vulnerability or any number of factors
#ac/vulnerability and amount of income gained outside of fishing
#ac/vulnerability and role in the industry, etc

ac_barnes$age = responses$age

age_ac<-ggboxplot(ac_barnes, x = "age", y = "new_ac",
          color = "age", palette = c("#00AFBB", "#E7B800", "#FC4E07", "#00AFBB", "#E7B800", "#FC4E07"),
          order = c("U30", "30-40", "40-50", "50-60", "60-70", "70+"),
          ylab = "adaptive capacity", xlab = "age")

ggsave(plot = age_ac, file = paste0("../figures/ac/age_ac.png"))
age_ac

#ac
#anova
age_aov<-aov(new_ac ~ age, data = ac_barnes)
summary(age_aov)

#tukey HSD of anova
age_tukey<-TukeyHSD(age_aov)
age_tukey
```

##anova for adaptive capacity by vessel length
```{r echo=FALSE, warning=FALSE}
#can copy and paste to look at age and vulnerability or any number of factors
#ac/vulnerability and amount of income gained outside of fishing
#ac/vulnerability and role in the industry, etc

ac_barnes$vessel_length = responses$vessel

vl_ac<-ggboxplot(ac_barnes, x = "vessel_length", y = "new_ac",
          color = "vessel_length", palette = c("#00AFBB", "#E7B800", "#FC4E07", "#00AFBB", "#E7B800", "#FC4E07"),
          order = c("U25", "26-35", "36-45", "46-55", "56-65", "66+"),
          ylab = "adaptive capacity", xlab = "vessel length")

ggsave(plot = age_ac, file = paste0("../figures/ac/age_ac.png"))
vl_ac

#ac
#anova
vl_aov<-aov(new_ac ~ vessel_length, data = ac_barnes)
summary(vl_aov)

#tukey HSD of anova
vl_tukey<-TukeyHSD(vl_aov)
vl_tukey
```
Generating a list of fisheries like in vulnerability, but here only fisheries that contribute at least 25% of income
#subseting by fishery
```{r make a new column and specify fishery(ies) for each person, echo=FALSE}

#modifed from code written by Laura Koehn
#use to calculate average vulnerability across a fishery
#modify for each survey

responses$fisheries25 = rep(NA, length =  nrow(responses))
# loop through each person and get what they are fishing for, put in new column
# should list each fishery they fish for
# not beautiful code but works
for(i in 1: nrow(responses)) {
  fisheriesname = ""
  if(responses$CPS_income[i] == 1) {
    fisheriesname = paste(fisheriesname, "CPS", sep = "") 
  }
  if(responses$dungeness_crab_income[i] == 1) {
    if(fisheriesname == "") { # have to do extra to not put a , at the beginning
      # if this is the first fishery 
      fisheriesname = paste(fisheriesname, "dungeness crab", sep = "")
    } else {
      fisheriesname = paste(fisheriesname, "dungeness crab", sep = ", ")
    }
  }
  if(responses$groundfish_income[i] == 1) {
    if(fisheriesname == "") {
      fisheriesname = paste(fisheriesname, "groundfish", sep = "")
    } else {
      fisheriesname = paste(fisheriesname, "groundfish", sep = ", ")
    }
  }
  if(responses$hake_income[i] == 1) {
    if(fisheriesname == "") {
    fisheriesname = paste(fisheriesname, "hake", sep = "")
    } else {
      fisheriesname = paste(fisheriesname, "hake", sep = ", ")
    }
  }
  if(responses$HMS_income[i] == 1) {
    if(fisheriesname == "") {
      fisheriesname = paste(fisheriesname, "HMS", sep = "")
    } else {
      fisheriesname = paste(fisheriesname, "HMS", sep = ", ")
    }
  }
  if(responses$salmon_income[i] == 1) {
    if(fisheriesname == "") {
      fisheriesname = paste(fisheriesname, "salmon", sep = "")
    } else {
      fisheriesname = paste(fisheriesname, "salmon", sep = ", ")
    }
  }
  if(responses$scallops_income[i] == 1) {
    if(fisheriesname == "") {
      fisheriesname = paste(fisheriesname, "scallops", sep = "")
    } else {
      fisheriesname = paste(fisheriesname, "scallops", sep = ", ")
    }
  }
  if(responses$urchin_income[i] == 1) {
    if(fisheriesname == "") {
      fisheriesname = paste(fisheriesname, "sea urchin", sep = "")
    } else {
      fisheriesname = paste(fisheriesname, "sea urchin", sep = ", ")
    }
  }
  if(responses$shrimp_income[i] == 1) {
    if(fisheriesname == "") {
      fisheriesname = paste(fisheriesname, "shrimp", sep = "")
    } else {
      fisheriesname = paste(fisheriesname, "shrimp", sep = ", ")
    }
  }
  if(responses$squid_income[i] == 1) {
    if(fisheriesname == "") {
      fisheriesname = paste(fisheriesname, "squid", sep = "")
    } else {
      fisheriesname = paste(fisheriesname, "squid", sep = ", ")
    }
  }
  if(responses$other_income[i] == 1) {
    if(fisheriesname == "") { 
    fisheriesname = paste(fisheriesname, "other", sep = "")
    #paste in what the person put for "other"
    }else {
      fisheriesname = paste(fisheriesname, "other", sep = ", ")
    }
  }
  responses$fisheries25[i] = fisheriesname
}

### Average across each UNIQUE "fishery" - so if they fish for salmon AND HMS, gets
#treated different than just fishing for salmon
#could use if fuction to pull out whatever "fishery" is desired with this method

fisheryavg = responses %>%
  group_by(fisheries25) %>%
  dplyr::summarise(exposure_avg = mean(indv_exposure, na.rm = TRUE),
                   sensitivity_avg = mean(indv_sensitivity, na.rm = TRUE),
                   adaptive_avg = mean(indv_ac),
                   vuln_avg = mean(indv_vulnerability),
                   vuln_euc_avg = mean(indv_vulnerability_euc, na.rm = TRUE),
                   count = n()
)

fisheryavg

#save too big to show up in rmd
write.csv(fisheryavg, paste0("../data/vulnerability_fisherycombos_25.csv"))
```