---
title: "data exploration_PNW"
author: "Jocelyn Runnebaum with edits from Laura Nelson"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
   word_document:
    reference_docx: word-styles-reference.docx

---

```{r setup, echo = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message = FALSE, error=FALSE)
```

```{r, packages}
library(tidyverse)
library(lubridate)
library(knitr)
library(grDevices)
library(extrafont)
library(ggrepel)
library(ggthemes)
# windowsFonts(Times=windowsFont("Calibri"))
theme_sleek <- function(base_size = 12, base_family = "Calibri") {
  half_line <- base_size/2
  theme_light(base_size = 12, base_family = "Calibri") +
    theme(
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      axis.ticks.length = unit(half_line / 2.2, "pt"),
      strip.background = element_rect(fill = NA, colour = NA),
      strip.text.x = element_text(colour = "black"),
      strip.text.y = element_text(colour = "black"),
      # axis.text = element_text(colour = "grey30"),
      #axis.title = element_text(colour = "grey30"),
      #legend.title = element_text(colour = "grey30"),#, size = rel(0.9)
      legend.title = element_blank(),
      panel.border = element_rect(fill = NA),#, colour = "grey70", size = 1),
      legend.key.size = unit(0.9, "lines"),
      #legend.text = element_text(size = rel(0.7)),#, colour = "grey30"),
      legend.key = element_rect(colour = NA, fill = NA),
      legend.background = element_rect(colour = NA, fill = NA)#,
      #plot.title = element_text(colour = "grey30"),#, size = rel(1),
      #plot.subtitle = element_text(colour = "grey30")#, size = rel(.85),
     
      
    )
}
theme_set(theme_sleek())
options(scipen = 999) #turns off scientific notation
# Depends on dplyr
tickr <- function(
  data, # dataframe
  var, # column of interest
  to # break point definition 
){
  
  VAR <- enquo(var) # makes VAR a dynamic variable
  
  data %>% 
    distinct(!!VAR) %>%
    ungroup(!!VAR) %>% 
    mutate(labels = ifelse(!!VAR %in% seq(to * round(min(!!VAR) / to), max(!!VAR), to),
                           !!VAR, "")) %>%
    select(breaks = UQ(VAR), labels)
}
```

```{r load data}
responses <- read.csv("../data/PNW_decoded_responses_30.Nov.2020.csv")

#add leading zeros back to zip codes in responses dataframe (they drop off when saving from data cleaning)
#should not be a problem for west coast zips but leaving in just in case

#for(i in 1:length(responses$home_zip)){
  #if(as.numeric(responses$home_zip[i]) < 10000){
    #responses$home_zip[i] <- paste0("0", responses$home_zip[i])
  #}
#}

#for(i in 1:length(responses$port_zip)){
  #if(as.numeric(responses$port_zip[i]) < 10000){
    #responses$port_zip[i] <- paste0("0", responses$port_zip[i])
  #}
#}
```

```{r zipcode data}
#download zipcode data to match states to responses data by zip code (could add county or other identifying information from these data)

ZipCodeSourceFile = "http://download.geonames.org/export/zip/US.zip"
temp <- tempfile()
download.file(ZipCodeSourceFile , temp)
ZipCodes <- read.table(unz(temp, "US.txt"), sep="\t")
unlink(temp)
names(ZipCodes) = c("CountryCode", "home_zip", "PlaceName", 
                    "state", "state_code", "County", "AdminCode2", 
                    "AdminName3", "AdminCode3", "latitude", "longitude", "accuracy")

zips<- ZipCodes %>% 
  select(home_zip, state, state_code) %>% 
  mutate(home_zip = as.integer(home_zip))

portzips<-ZipCodes %>% 
  select(home_zip, state, state_code) %>% 
  mutate(home_zip = as.integer(home_zip))

names(portzips) = c("port_zip", "port_state", "port_state_code")

#add leading zero back
#for(i in 1:length(zips$home_zip)){
  #if(as.numeric(zips$home_zip[i]) < 10000){
    #zips$home_zip[i] <- paste0("0", zips$home_zip[i])
  #}
#}
```

```{r add state and port data to responses}

responses <- responses %>% 
  left_join(zips)

responses <- responses %>% 
  left_join(portzips)

belief <- responses %>% 
  select(age, state, state_code, ability, believe, harm_me, harm_future)

```

```{r save data}
write.csv(responses, paste0("../data/PNW_decoded_responses_wstate", format(Sys.time(), "%d.%b.%Y"),".csv"))
```
#Responses by State
```{r responses by State}
#update based upon survey region
responses$state <- factor(responses$state, levels = c("Washington", "Oregon", "California", "Alaska"))

responses_by_state_plot <- ggplot(data = responses, aes(x=state))+
  geom_bar(stat = "count")+
  xlab("State\n")+
  ggtitle("Respondents by State") 

ggsave(plot = responses_by_state_plot, file = paste0("../figures/resp_sum/response_by_state.png"))

#percentage instead of count
responses_by_state_plot <- ggplot(data = responses, aes(x=state))+
  geom_bar(aes(y = (..count..)/sum(..count..)), fill = c("light blue", "navy blue", "purple", "forest green"))+
  scale_y_continuous(labels=scales::percent) +
  xlab("State\n")+ ylab("percent") +
  ggtitle("Respondents by State") 

ggsave(plot = responses_by_state_plot, file = paste0("../figures/resp_sum/percent_response_by_state.png"))

```

#basic info summary
```{r responses for q 5 - 10}
#q5 how long have you been fishing
responses$yrs_fishing<-factor(responses$yrs_fishing, levels = c("0-5", "5-15", "15-25","25+"))

yrsfishing_plot <- ggplot(data = responses, aes (x=yrs_fishing))+
  geom_bar(fill = "blue", position = position_dodge(), stat="count")+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1, size = 16))+ ylab("Number of Respondents\n")+
  xlab("Category\n")+
  ggtitle("How long have you been fishing") 
ggsave(plot = yrsfishing_plot, file = paste0("../figures/resp_sum/yrsfishing.png"))

#percentage instead of count

#q8 what is your role in the fishing industry

#q9 how many crew do you employ

#q10 what length vessel do you work on or operate
responses$vessel <- factor(responses$vessel, levels = c("U25", "26-35", "36-45", "46-55", "56-65", "66+"))

vessel_plot <- ggplot(data = responses, aes(x=vessel))+
  geom_bar(stat = "count")+
  xlab("vessel\n")+
  ggtitle("Respondents by Vessel length") 

ggsave(plot = vessel_plot, file = paste0("../figures/resp_sum/responses_vessel.png"))
```

#gender, age, income, tribe demographic summary
```{r responses for q12 - 17}
#q12 Do you currently participate in fisheries in other parts of the country?
responses$other_fisheries<-factor(responses$other_fisheries, levels = c("yes", "no"))

otherfisheries_plot <- ggplot(data = responses, aes (x=other_fisheries))+
  geom_bar(fill = "blue", position = position_dodge(), stat="count")+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1, size = 16))+ ylab("Number of Respondents\n")+
  xlab("Category\n")+
  ggtitle("Do you participate in fisheries in other parts of the country") 

ggsave(plot = otherfisheries_plot, file = paste0("../figures/resp_sum/other_fisheries.png"))

#q13 is an open ended response

#q14 how much income comes from outside of fishing
responses$income <- factor(responses$income, levels = c("none", "U10", "10-25", "25-50", "50+"))

responses_income_plot <- ggplot(data = responses, aes(x=income))+
  geom_bar(stat = "count")+
  xlab("percent of income")+
  ggtitle("Percentage of income from outside fishing") 

ggsave(plot = responses_income_plot, file = paste0("../figures/resp_sum/responses_income.png"))

#q15 how old are you
responses$age <- factor(responses$age, levels = c("U30", "30-40", "40-50", "50-60", "60-70", "70+"))

responses_by_age_plot <- ggplot(data = responses, aes(x=age))+
  geom_bar(position = position_dodge(),stat = "count")
ggsave(plot = responses_by_age_plot, file = paste0("../figures/resp_sum/responses_age.png"))

#q16 What is your gender identity?
responses$gender <- factor(responses$gender, levels = c("female", "male", "no_response", "other"))

responses_gender_plot <- ggplot(data = responses, aes(x=gender))+
  geom_bar(position = position_dodge(),stat = "count")
ggsave(plot = responses_gender_plot, file = paste0("../figures/resp_sum/responses_gender.png"))

#q17 are you an enrolled tribal member?
responses$tribe<-factor(responses$tribe, levels = c("yes", "no"))

tribe_plot <- ggplot(data = responses, aes (x=tribe))+
  geom_bar(fill = "blue", position = position_dodge(), stat="count")+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1, size = 16))+ ylab("Number of Respondents\n")+
  xlab("Category\n")+
  ggtitle("Are you an enrolled tribal member?") 

ggsave(plot = tribe_plot, file = paste0("../figures/resp_sum/tribe.png"))
```
#observations of ocean change
```{r q20 observations of change}
#changes you have observed in the last five years?
obs <- responses %>% 
  select(age, ocean_temp, severe_weather, target_spp)

#q20r1 changes in ocean temperature
obs$ocean_temp <- factor(obs$ocean_temp, levels = c("no_change", "increase", "decrease"))

oceanobs_plot <- ggplot(data = obs, aes (x=ocean_temp))+
  geom_bar(fill = "blue", position = position_dodge(), stat="count")+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1, size = 16))+ ylab("Number of Respondents\n")+
  xlab("Category\n")+
  ggtitle("Observations of change in ocean temperature") 

ggsave(plot = oceanobs_plot, file = paste0("../figures/ocean_temp_obs.png"))

#q20r2 changes in severe weather
obs$severe_weather <- factor(obs$severe_weather, levels = c("no_change", "increase", "decrease"))

weatherobs_plot <- ggplot(data = obs, aes (x=severe_weather))+
  geom_bar(fill = "blue", position = position_dodge(), stat="count")+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1, size = 16))+ ylab("Number of Respondents\n")+
  xlab("Category\n")+
  ggtitle("Observations of change in severe weather") 

ggsave(plot = weatherobs_plot, file = paste0("../figures/weather_obs.png"))

#q20r3 changes in target species
obs$target_spp <- factor(obs$target_spp, levels = c("no_change", "increase", "decrease"))

targetspp_plot <- ggplot(data = obs, aes (x=target_spp))+
  geom_bar(fill = "blue", position = position_dodge(), stat="count")+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1, size = 16))+ ylab("Number of Respondents\n")+
  xlab("Category\n")+
  ggtitle("Observations of change in target species") 

ggsave(plot = targetspp_plot, file = paste0("../figures/targetspp_obs.png"))
```
##ability to catch fish has been affected by climate change
```{r ability to catch fish}
#q21 Do you think your ability to catch fish has been affected by climate change?
#This is an interesting plot, it doesn't appear people think that climate change has really impacted their ability to catch fish
#does this mean they think their fishing ability is indepent of climate change?
belief$ability <- factor(belief$ability, levels = c("no", "no_obs_change", "yes_neg", "yes_pos"))

cc_ability_plot <- ggplot(data = belief, aes (x=ability))+
  geom_bar(position = position_dodge(), stat="count")+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))+ ylab("Number of Respondents\n")+
  xlab("Category\n")+
  ggtitle("Climate change has affected ability to catch fish") 

ggsave(plot = cc_ability_plot, file = paste0("../figures/abilitytofish.png"))
```
#see climate and fisheries code for q23-24


##Belief in climate change by State
```{r belief in climate change by State}

belief$believe <- factor(belief$believe, levels = c("strongly_disagree", "somewhat_disagree", "neutral", "somewhat_agree", "strongly_agree"))

belief$state <- factor(belief$state, levels = c("Washington","Oregon", "California", "Alaska"))

cc_belief_state_plot <- ggplot(data = belief, aes (x=believe, fill = state))+
  geom_bar(position = position_dodge(), stat="count")+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))+ ylab("Number of Respondents\n")+
  xlab("Category\n")+
  ggtitle("Believe that climate change is occuring") 

ggsave(plot = cc_belief_state_plot, file = paste0("../figures/climate_change_belief_state.png"))

```

##Believe that climate change will harm them by State
```{r climate change harm me - state}
#Believe that climate change will harm them personally
belief$harm_me <- factor(belief$harm_me, levels = c("strongly_disagree", "somewhat_disagree", "neutral", "somewhat_agree", "strongly_agree"))

belief$state <- factor(belief$state, levels = c("Washington", "Oregon", "California", "Alaska"))

cc_harm_me_by_state_plot <- ggplot(data = belief, aes (x=harm_me, fill = state))+
  geom_bar(position = position_dodge(), stat="count")+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))+ ylab("Number of Respondents\n")+
  xlab("Category\n")+
  ggtitle("Believe that climate change will harm them personally") 

ggsave(plot = cc_harm_me_by_state_plot, file = paste0("../figures/climate_change_harm_me_state.png"))

```

##Believe that climate change will harm future generations by State
```{r climate change harm future - state}
#Believe that climate change will harm future generations
belief$harm_future <- factor(belief$harm_future, levels = c("strongly_disagree", "somewhat_disagree", "neutral", "somewhat_agree", "strongly_agree"))

belief$state <- factor(belief$state, levels = c("Washington", "Oregon", "California", "Alaska"))

cc_harm_future_by_state_plot <- ggplot(data = belief, aes (x=harm_future, fill = state))+
  geom_bar(position = position_dodge(), stat="count")+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))+ ylab("Number of Respondents\n")+
  xlab("Category\n")+
  ggtitle("Believe that climate change will harm future generations") 

ggsave(plot = cc_harm_future_by_state_plot, file = paste0("../figures/climate_change_harm_future_state.png"))

```

##Belief in climate change by age
```{r belief in climate change}
belief$believe <- factor(belief$believe, levels = c("strongly_disagree", "somewhat_disagree", "neutral", "somewhat_agree", "strongly_agree"))
belief$age <- factor(belief$age, levels = c("U30", "30-40", "40-50", "50-60", "60-70", "70+"))


cc_belief_plot <- ggplot(data = belief, aes (x=believe, fill = age))+
  geom_bar(position = position_dodge(), stat="count")+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))+ ylab("Number of Respondents\n")+
  xlab("Category\n")+
  ggtitle("Believe that climate change is occuring") 

ggsave(plot = cc_belief_plot, file = paste0("../figures/climate_change_belief_age.png"))

```

##Believe that climate change will harm them by age
```{r climate change harm me - age}
#Believe that climate change will harm them personally
belief$harm_me <- factor(belief$harm_me, levels = c("strongly_disagree", "somewhat_disagree", "neutral", "somewhat_agree", "strongly_agree"))

belief$age <- factor(belief$age, levels = c("U30", "30-40", "40-50", "50-60", "60-70", "70+"))

cc_harm_me_by_age_plot <- ggplot(data = belief, aes (x=harm_me, fill = age))+
  geom_bar(position = position_dodge(), stat="count")+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))+ ylab("Number of Respondents\n")+
  xlab("Category\n")+
  ggtitle("Believe that climate change will harm them personally") 

ggsave(plot = cc_harm_me_by_age_plot, file = paste0("../figures/climate_change_harm_me_age.png"))

```

##Believe that climate change will harm future generations by age
```{r climate change harm future - age}
#Believe that climate change will harm future generations
belief$harm_future <- factor(belief$harm_future, levels = c("strongly_disagree", "somewhat_disagree", "neutral", "somewhat_agree", "strongly_agree"))

belief$age <- factor(belief$age, levels = c("U30", "30-40", "40-50", "50-60", "60-70", "70+"))

cc_harm_future_by_age_plot <- ggplot(data = belief, aes (x=harm_future, fill = age))+
  geom_bar(position = position_dodge(), stat="count")+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))+ ylab("Number of Respondents\n")+
  xlab("Category\n")+
  ggtitle("Believe that climate change will harm future generations") 

ggsave(plot = cc_harm_future_by_age_plot, file = paste0("../figures/climate_change_harm_future_age.png"))

```


##ability to catch fish has been affected by climate change
```{r ability to catch fish}
#Do you think your ability to catch fish has been affected by climate change?
#This is an interesting plot, it doesn't appear people think that climate change has really impacted their ability to catch fish
#does this mean they think their fishing ability is indepent of climate change?
belief$ability <- factor(belief$ability, levels = c("no", "no_obs_change", "yes_neg", "yes_pos"))

belief$age <- factor(belief$age, levels = c("U30", "30-40", "40-50", "50-60", "60-70", "70+"))

cc_ability_by_age_plot <- ggplot(data = belief, aes (x=ability, fill = age))+
  geom_bar(position = position_dodge(), stat="count")+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))+ ylab("Number of Respondents\n")+
  xlab("Category\n")+
  ggtitle("Climate change has affected ability to catch fish") 

ggsave(plot = cc_ability_by_age_plot, file = paste0("../figures/climate_change_ability_age.png"))

```

