---
title: "data exploration"
author: "Jocelyn Runnebaum with edits from Laura Nelson"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
   word_document:
    reference_docx: word-styles-reference.docx

---

```{r setup, echo = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message = FALSE, error=FALSE)
```

```{r, packages}
library(tidyverse)
library(lubridate)
library(knitr)
library(grDevices)
library(extrafont)
library(ggrepel)
library(ggthemes)
library(PNWColors)
library(ggplot2)
library(gridExtra)
# windowsFonts(Times=windowsFont("Calibri"))
theme_sleek <- function(base_size = 12, base_family = "Calibri") {
  half_line <- base_size/2
  theme_light(base_size = 12, base_family = "Calibri") +
    theme(
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      axis.ticks.length = unit(half_line / 2.2, "pt"),
      strip.background = element_rect(fill = NA, colour = NA),
      strip.text.x = element_text(colour = "black"),
      strip.text.y = element_text(colour = "black"),
      # axis.text = element_text(colour = "grey30"),
      #axis.title = element_text(colour = "grey30"),
      #legend.title = element_text(colour = "grey30"),#, size = rel(0.9)
      legend.title = element_blank(),
      panel.border = element_rect(fill = NA),#, colour = "grey70", size = 1),
      legend.key.size = unit(0.9, "lines"),
      #legend.text = element_text(size = rel(0.7)),#, colour = "grey30"),
      legend.key = element_rect(colour = NA, fill = NA),
      legend.background = element_rect(colour = NA, fill = NA)#,
      #plot.title = element_text(colour = "grey30"),#, size = rel(1),
      #plot.subtitle = element_text(colour = "grey30")#, size = rel(.85),
     
      
    )
}
theme_set(theme_sleek())
options(scipen = 999) #turns off scientific notation
# Depends on dplyr
tickr <- function(
  data, # dataframe
  var, # column of interest
  to # break point definition 
){
  
  VAR <- enquo(var) # makes VAR a dynamic variable
  
  data %>% 
    distinct(!!VAR) %>%
    ungroup(!!VAR) %>% 
    mutate(labels = ifelse(!!VAR %in% seq(to * round(min(!!VAR) / to), max(!!VAR), to),
                           !!VAR, "")) %>%
    select(breaks = UQ(VAR), labels)
}
```

```{r load data}
#change to appropriate file name

responses <- read.csv("../data/PNW_decoded_responses_27.Jul.2021.csv")
```

#add leading zeros back to zip codes in responses dataframe (they drop off when saving from data cleaning)
#only needed for new england survey, do not run for other versions
```{r}

for(i in 1:length(responses$home_zip)){
  if(as.numeric(responses$home_zip[i]) < 10000){
    responses$home_zip[i] <- paste0("0", responses$home_zip[i])
  }
}

for(i in 1:length(responses$port_zip)){
  if(as.numeric(responses$port_zip[i]) < 10000){
    responses$port_zip[i] <- paste0("0", responses$port_zip[i])
  }
}
```

```{r zipcode data}
#download zipcode data to match states to responses data by zip code (could add county or other identifying information from these data)
#for survey versions in the US

ZipCodeSourceFile = "http://download.geonames.org/export/zip/US.zip"
temp <- tempfile()
download.file(ZipCodeSourceFile , temp)
ZipCodes <- read.table(unz(temp, "US.txt"), sep="\t")
unlink(temp)
names(ZipCodes) = c("CountryCode", "home_zip", "PlaceName", 
                    "state", "state_code", "County", "AdminCode2", 
                    "AdminName3", "AdminCode3", "latitude", "longitude", "accuracy")

zips<- ZipCodes %>% 
  select(home_zip, state, state_code) %>% 
  mutate(home_zip = as.integer(home_zip))

portzips<-ZipCodes %>% 
  select(home_zip, state, state_code) %>% 
  mutate(home_zip = as.integer(home_zip))

names(portzips) = c("port_zip", "port_state", "port_state_code")

#add leading zero back
#hash out for versions of survey where this is not needed
#for(i in 1:length(zips$home_zip)){
  #if(as.numeric(zips$home_zip[i]) < 10000){
    #zips$home_zip[i] <- paste0("0", zips$home_zip[i])
  #}
#}

#for(i in 1:length(portzips$port_zip)){
  #if(as.numeric(portzips$port_zip[i]) < 10000){
    #portzips$port_zip[i] <- paste0("0", portzips$port_zip[i])
  #}
#}
```

#adds columns for state and state code for residence and homeport
#state and state_code are residental
#port_state and port_state_code are homeport
```{r add state and port data to responses}

responses <- responses %>% 
  left_join(zips)

responses <- responses %>% 
  left_join(portzips)

```


##This section produces basic summaries of the survey responses including demographics and numbers of responses.
#saves plots to a folder in figures called resp_sum (responses summary)

#Responses by State
#update by survey region
```{r responses by State}

#change to correct states
responses$state <- factor(responses$state, levels = c("Washington", "Oregon", "California"))

responses_by_state_plot <- ggplot(data = responses, aes(x=state))+
  geom_bar(stat = "count")+
  xlab("State\n")+
  ggtitle("Respondents by State") 

ggsave(plot = responses_by_state_plot, file = paste0("../figures/resp_sum/response_by_state.png"))

#percentage of responses instead of count
responses_by_state_plot <- ggplot(data = responses, aes(x=state))+
  geom_bar(aes(y = (..count..)/sum(..count..)), fill = "#749dae") +
  scale_y_continuous(labels=scales::percent) +
  xlab("State\n")+ ylab("percent") +
  ggtitle("Respondents by State") 

ggsave(plot = responses_by_state_plot, file = paste0("../figures/resp_sum/percent_response_by_state.png"))

```

#color palettes by how many categories there are
```{r}
pal3<-pnw_palette("Sailboat", 3, type = "discrete")
pal<-pnw_palette("Sailboat", 4, type = "discrete")
pal5<-pnw_palette("Sailboat", 5, type = "discrete")
pal6<-pnw_palette("Sailboat", 6, type = "discrete")
```

#basic info summary
#update by survey region
```{r responses for q 3 - 10}
#q3 regions where people fish
#put in correct regions

regions <-  responses %>%
  select(pugetsound_SJF, wa_coast, cr, or_coast, norcal, cencal, socal) %>%  #other_region is also an option
  gather(region, response) %>% 
  group_by(region)
regions$region <- factor(regions$region, levels = c("pugetsound_SJF", "wa_coast", "cr", "or_coast", "norcal", "cencal", "socal"))
region_plot <- ggplot()+
  geom_col(data = regions, aes(x=region, y = response), fill = "#749dae")+
  xlab("Species\n")+
  ylab("Count\n")+
  ggtitle("Regions Fished\n")

ggsave(plot = region_plot, file = paste0("../figures/resp_sum/regions_fished.png"))

#q4 is an open ended question asking where is your homeport

#q5 how long have you been fishing
responses$yrs_fishing<-factor(responses$yrs_fishing, levels = c("0-5", "5-15", "15-25","25+"))

yrsfishing_plot <- ggplot(data = responses, aes (x=yrs_fishing))+
  geom_bar(fill = pal, position = position_dodge(), stat="count")+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1, size = 16))+ ylab("Number of Respondents\n")+
  xlab("Category\n")+
  ggtitle("How long have you been fishing") 
ggsave(plot = yrsfishing_plot, file = paste0("../figures/resp_sum/yrsfishing.png"))

#percentage of responses instead of count
yrs_fishing_pct <- ggplot(data = responses, aes(x=yrs_fishing))+
  geom_bar(aes(y = (..count..)/sum(..count..)), fill = pal)+
  scale_y_continuous(labels=scales::percent) +
  xlab("yrs fishing\n")+ ylab("percent") +
  ggtitle("Years Fishing") 

ggsave(plot = yrs_fishing_pct, file = paste0("../figures/resp_sum/yrsfishing_pct.png"))

#q6 is an open ended question asking where you live

#q7 how long have you lived in your current community
responses$yrs_residence<-factor(responses$yrs_residence, levels = c("0-5", "5-15", "15-25","25+"))

yrsresidence_plot <- ggplot(data = responses, aes (x=yrs_residence))+
  geom_bar(fill = pal, position = position_dodge(), stat="count")+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1, size = 16))+ ylab("Number of Respondents\n")+
  xlab("Category\n")+
  ggtitle("How long have you lived in your community") 
ggsave(plot = yrsresidence_plot, file = paste0("../figures/resp_sum/yrsresidence.png"))

#percentage of responses instead of count
yrs_res_pct <- ggplot(data = responses, aes(x=yrs_residence))+
  geom_bar(aes(y = (..count..)/sum(..count..)), fill = pal)+
  scale_y_continuous(labels=scales::percent) +
  xlab("yrs residence\n")+ ylab("percent") +
  ggtitle("Years Living in Community") 

ggsave(plot = yrs_res_pct, file = paste0("../figures/resp_sum/yrsres_pct.png"))

#q8 what is your role in the fishing industry
job <-  responses %>%
  select(captain, vessel_owner, crew, other_role) %>%  
  gather(role, response) %>% 
  group_by(role)
job$role <- factor(job$role, levels = c("captain", "vessel_owner", "crew", "other_role"))
job_plot <- ggplot()+
  geom_col(data = job, aes(x=role, y = response), fill = "#749dae")+
  xlab("Role\n")+
  ylab("Count\n")+
  ggtitle("Industry role\n")

ggsave(plot = job_plot, file = paste0("../figures/resp_sum/industry_role.png"))

#q9 how many crew do you employ
responses$crew_num<-factor(responses$crew_num, levels = c("1-3", "4-7", "7-11","11+"))

crewemployed_plot <- ggplot(data = responses, aes (x=crew_num))+
  geom_bar(fill = "navy", position = position_dodge(), stat="count")+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1, size = 16))+ ylab("Number of Respondents\n")+
  xlab("Category\n")+
  ggtitle("How many crew do you employ") 
ggsave(plot = crewemployed_plot, file = paste0("../figures/resp_sum/crew_employed.png"))

#q10 what length vessel do you work on or operate
#may need to be modified by region
responses$vessel <- factor(responses$vessel, levels = c("U25", "26-35", "36-45", "46-55", "56-65", "66+"))

vessel_plot <- ggplot(data = responses, aes(x=vessel))+
  geom_bar(stat = "count")+
  xlab("vessel\n")+
  ggtitle("Respondents by Vessel length") 

ggsave(plot = vessel_plot, file = paste0("../figures/resp_sum/responses_vessel.png"))

#percentage instead of count
vessel_pct <- ggplot(data = responses, aes(x=vessel))+
  geom_bar(aes(y = (..count..)/sum(..count..)), fill = pal6)+
  scale_y_continuous(labels=scales::percent) +
  xlab("vessel length\n")+ ylab("percent") +
  ggtitle("What length vessel do you work on") 

ggsave(plot = vessel_pct, file = paste0("../figures/resp_sum/vessel_pct.png"))


#q11 what gear types do you use
#needs to be updated by region
gears <-  responses %>%
  select(longline, midwater_trwl, bottom_trwl, pots, purse_seine, gillnet, troll, other_gear) %>%  
  gather(gt, response) %>% 
  group_by(gt)
gears$gt <- factor(gears$gt, levels = c("longline", "midwater_trwl", "bottom_trwl", "pots", "purse_seine", "gillnet", "troll", "other_gear"))
gear_plot <- ggplot()+
  geom_col(data = gears, aes(x=gt, y = response), fill = "#749dae")+
  xlab("Gear type\n")+
  ylab("Count\n")+
  ggtitle("Gear type\n")

ggsave(plot = gear_plot, file = paste0("../figures/resp_sum/gear_type.png"))
```

##q 8 and 11 are in long format
##this adds a column at the end of the sheet with the responses so summarize can be used to see differences between the groups

#q8 industry role
#can add in details from open ended responses about other jobs if desired
```{r}

responses$industry_role = rep(NA, length =  nrow(responses))
for(i in 1: nrow(responses)) {
  inrole = ""
  if(responses$captain[i] == 1) {
    inrole = paste(inrole, "captain", sep = "") 
  }
  if(responses$vessel_owner[i] == 1) {
    if(inrole == "") { # have to do extra to not put a , at the beginning
      # if this is the first fishery 
      inrole = paste(inrole, "vessel owner", sep = "")
    } else {
      inrole = paste(inrole, "vessel owner", sep = ", ")
    }
  }
  if(responses$crew[i] == 1) {
    if(inrole == "") {
      inrole = paste(inrole, "crew", sep = "")
    } else {
      inrole = paste(inrole, "crew", sep = ", ")
    }
  }
  if(responses$other_role[i] == 1) {
    if(inrole == "") { 
    inrole = paste(inrole, "other", sep = "")
    #just use "other" for fewer categories
    }else {
      inrole = paste(inrole, "other", sep = ", ")
    }
  }
  responses$industry_role[i] = inrole
}
#add if details wanted about other role  
#if(responses$other_role[i] == 1) {
    #if(inrole == "") { 
    #inrole = paste(inrole, responses$other_role_descrip[i], sep = "")
    #paste in what the person put for "other"
    #}else {
      #inrole = paste(inrole, responses$other_role_descrip[i], sep = ", ")
    #}
  #}
  #responses$industry_role[i] = inrole
#}
```

#q11 gear types
#can add in details from open ended responses about other gear types if desired
```{r}

responses$gear_type = rep(NA, length =  nrow(responses))
for(i in 1: nrow(responses)) {
  gtype = ""
  if(responses$longline[i] == 1) {
    gtype = paste(gtype, "longline", sep = "") 
  }
  if(responses$midwater_trwl[i] == 1) {
    if(gtype == "") {  
      gtype = paste(gtype, "midwater trawl", sep = "")
    } else {
      gtype = paste(gtype, "midwater trawl", sep = ", ")
    }
  }
  if(responses$bottom_trwl[i] == 1) {
    if(gtype == "") {  
      gtype = paste(gtype, "bottom trawl", sep = "")
    } else {
      gtype = paste(gtype, "bottom trawl", sep = ", ")
    }
  }
  if(responses$pots[i] == 1) {
    if(gtype == "") {  
      gtype = paste(gtype, "pots", sep = "")
    } else {
      gtype = paste(gtype, "pots", sep = ", ")
    }
  }
  if(responses$purse_seine[i] == 1) {
    if(gtype == "") {  
      gtype = paste(gtype, "purse seine", sep = "")
    } else {
      gtype = paste(gtype, "purse seine", sep = ", ")
    }
  }
  if(responses$gillnet[i] == 1) {
    if(gtype == "") {  
      gtype = paste(gtype, "gillnet", sep = "")
    } else {
      gtype = paste(gtype, "gillnet", sep = ", ")
    }
  }
  if(responses$troll[i] == 1) {
    if(gtype == "") {  
      gtype = paste(gtype, "troll", sep = "")
    } else {
      gtype = paste(gtype, "troll", sep = ", ")
    }
  }
  if(responses$other_gear[i] == 1) {
    if(gtype == "") { 
    gtype = paste(gtype, "other", sep = "")
    #just use "other" for fewer categories
    }else {
      inrole = paste(gtype, "other", sep = ", ")
    }
  }
  responses$gear_type[i] = gtype
}
#add if details wanted about other role  
#if(responses$other_gear[i] == 1) {
    #if(gtype == "") { 
    #gtype = paste(gtype, responses$other_gear_descrip[i], sep = "")
    #paste in what the person put for "other"
    #}else {
      #gtype = paste(gtype, responses$other_gear_descrip[i], sep = ", ")
    #}
  #}
  #responses$gear_type[i] = gtype
#}
```

#gender, age, income, tribe demographic summary
```{r responses for q12 - 17}
#q12 Do you currently participate in fisheries in other parts of the country?
responses$other_fisheries<-factor(responses$other_fisheries, levels = c("yes", "no"))

otherfisheries_plot <- ggplot(data = responses, aes (x=other_fisheries))+
 geom_bar(aes(y = (..count..)/sum(..count..)), fill = "#749dae")+
  scale_y_continuous(labels=scales::percent) +
  xlab("other fisheries\n")+ ylab("percent") +
  ggtitle("Do you participate in fisheries in other parts of the country") 

ggsave(plot = otherfisheries_plot, file = paste0("../figures/resp_sum/other_fisheries.png"))

#q13 is an open ended response asking what other regions if the previous answer was yes

#q14 how much income comes from outside of fishing
responses$income <- factor(responses$income, levels = c("none", "U10", "10-25", "25-50", "50+"))

responses_income_plot <- ggplot(data = responses, aes(x=income))+
  geom_bar(aes(y = (..count..)/sum(..count..)), fill = pal5)+
  scale_y_continuous(labels=scales::percent) +
  xlab("percent income\n")+ ylab("percent") +
  ggtitle("Percentage of income from outside fishing") 

ggsave(plot = responses_income_plot, file = paste0("../figures/resp_sum/responses_income.png"))

#q15 how old are you
responses$age <- factor(responses$age, levels = c("U30", "30-40", "40-50", "50-60", "60-70", "70+"))

responses_by_age_plot <- ggplot(data = responses, aes(x=age))+
  geom_bar(position = position_dodge(),stat = "count")
ggsave(plot = responses_by_age_plot, file = paste0("../figures/resp_sum/responses_age.png"))

#by percentage
ages_pct_plot <- ggplot(data = responses, aes(x=age))+
  geom_bar(aes(y = (..count..)/sum(..count..)), fill = pal6)+
  scale_y_continuous(labels=scales::percent) +
  xlab("age\n")+ ylab("percent") +
   ggtitle("Age groups")
ggsave(plot = ages_pct_plot, file = paste0("../figures/resp_sum/age_pct.png"))

#q16 What is your gender identity?
responses$gender <- factor(responses$gender, levels = c("female", "male", "no_response", "other"))

responses_gender_plot <- ggplot(data = responses, aes(x=gender))+
  geom_bar(position = position_dodge(),stat = "count")
ggsave(plot = responses_gender_plot, file = paste0("../figures/resp_sum/responses_gender.png"))

#q17 are you an enrolled tribal member?
responses$tribe<-factor(responses$tribe, levels = c("yes", "no"))

tribe_plot <- ggplot(data = responses, aes (x=tribe))+
  geom_bar(fill = "#749dae", position = position_dodge(), stat="count")+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1, size = 16))+ ylab("Number of Respondents\n")+
  xlab("Category\n")+
  ggtitle("Are you an enrolled tribal member?") 

ggsave(plot = tribe_plot, file = paste0("../figures/resp_sum/tribe.png"))
```

##smoking and drinking
```{r}
#q18 do you regularly smoke cigarettes
#saving these to wellbeing folder instead of general demographic summary location
responses$smoke<-factor(responses$smoke, levels = c("yes", "no"))

smoke_plot <- ggplot(data = responses, aes (x=smoke))+
  geom_bar(aes(y = (..count..)/sum(..count..)), fill = "navy blue")+
  scale_y_continuous(labels=scales::percent) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1, size = 16))+ ylab("Percent of Respondents\n")+
  ggtitle("Do you regularly smoke cigarettes?") 

ggsave(plot = smoke_plot, file = paste0("../figures/wellbeing/smoke.png"))

#q19 during the last 12 months on days when you drank alcohol many drinks did you usually have?

responses$alcohol<-factor(responses$alcohol, levels = c("1-2", "3-4", "5-6", "7-8", "8+", "NA", "prefer not to answer"))

drink_plot <- ggplot(data = responses, aes (x=alcohol))+
  geom_bar(aes(y = (..count..)/sum(..count..)), fill = "navy blue")+
  scale_y_continuous(labels=scales::percent) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1, size = 16))+ ylab("Percent of Respondents\n")+
  ggtitle("On days when you drink how many drinks did you usually have?") 

ggsave(plot = drink_plot, file = paste0("../figures/wellbeing/drink.png"))

```
#observations of ocean change
```{r q20 observations of change}
#changes you have observed in the last five years?
obs <- responses %>% 
  select(age, ocean_temp, severe_weather, target_spp)

#q20r1 changes in ocean temperature
obs$ocean_temp <- factor(obs$ocean_temp, levels = c("no_change", "increase", "decrease"))

oceanobs_plot <- ggplot(data = obs, aes (x=ocean_temp))+
  geom_bar(fill = "blue", position = position_dodge(), stat="count")+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1, size = 16))+ ylab("Number of Respondents\n")+
  xlab("Category\n")+
  ggtitle("Observations of change in ocean temperature") 

ggsave(plot = oceanobs_plot, file = paste0("../figures/resp_sum/ocean_temp_obs.png"))

#q20r2 changes in severe weather
obs$severe_weather <- factor(obs$severe_weather, levels = c("no_change", "increase", "decrease"))

weatherobs_plot <- ggplot(data = obs, aes (x=severe_weather))+
  geom_bar(fill = "blue", position = position_dodge(), stat="count")+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1, size = 16))+ ylab("Number of Respondents\n")+
  xlab("Category\n")+
  ggtitle("Observations of change in severe weather") 

ggsave(plot = weatherobs_plot, file = paste0("../figures/resp_sum/weather_obs.png"))

#q20r3 changes in target species
obs$target_spp <- factor(obs$target_spp, levels = c("no_change", "increase", "decrease"))

targetspp_plot <- ggplot(data = obs, aes (x=target_spp))+
  geom_bar(fill = "blue", position = position_dodge(), stat="count")+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1, size = 16))+ ylab("Number of Respondents\n")+
  xlab("Category\n")+
  ggtitle("Observations of change in target species") 

ggsave(plot = targetspp_plot, file = paste0("../figures/resp_sum/targetspp_obs.png"))
```

#observations of ocean change
#one figure and by percentages
```{r q20 observations of change}
#changes you have observed in the last five years?

#q20r1 changes in ocean temperature
oceanobs_plot <- ggplot(data = obs, aes (x=ocean_temp))+
  geom_bar(aes(y = (..count..)/sum(..count..)), fill = pal3)+
  scale_y_continuous(labels=scales::percent) +
  ylab("percent") +
  theme(axis.title.x = element_blank(), axis.text.x = element_text(angle = 90)) +
   ggtitle("Ocean temperature") 

#q20r2 changes in severe weather
weatherobs_plot <- ggplot(data = obs, aes (x=severe_weather))+
  geom_bar(aes(y = (..count..)/sum(..count..)), fill = pal3)+
  scale_y_continuous(labels=scales::percent) +
  theme(axis.title.x = element_blank(), axis.title.y = element_blank(), axis.text.x = element_text(angle = 90)) +
   ggtitle("Severe weather") 

#q20r3 changes in target species
targetspp_plot <- ggplot(data = obs, aes (x=target_spp))+
  geom_bar(aes(y = (..count..)/sum(..count..)), fill = pal3)+
  scale_y_continuous(labels=scales::percent) +
  theme(axis.title.x = element_blank(), axis.title.y = element_blank(), axis.text.x = element_text(angle = 90)) +
   ggtitle("Target species")

obs_plot<-grid.arrange(oceanobs_plot, weatherobs_plot, targetspp_plot, nrow = 1, top = "Have you observed changes to the following?")

ggsave(plot = obs_plot, file = paste0("../figures/species_change/ocean_obs.png"))

```
##ability to catch fish has been affected by climate change
```{r ability to catch fish}
#q21 Do you think your ability to catch fish has been affected by climate change?
#This is an interesting plot, it doesn't appear people think that climate change has really impacted their ability to catch fish
#does this mean they think their fishing ability is indepent of climate change?

belief <- responses %>% 
  select(age, state, state_code, ability, believe, harm_me, harm_future)
belief$ability <- factor(belief$ability, levels = c("no", "no_obs_change", "yes_neg", "yes_pos"))

cc_ability_plot <- ggplot(data = belief, aes (x=ability))+
  geom_bar(aes(y = (..count..)/sum(..count..)), fill = pal)+
  scale_y_continuous(labels=scales::percent) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))+ ylab("pct Respondents\n")+
  xlab("Category\n")+
  ggtitle("Climate change has affected ability to catch fish") 

ggsave(plot = cc_ability_plot, file = paste0("../figures/resp_sum/abilitytofish.png"))

#q22 is an open ended question about how their ability to fish has been affected by climate change
```

##ability to catch fish has been affected by climate change
```{r ability to catch fish}
#Do you think your ability to catch fish has been affected by climate change?
#This is an interesting plot, it doesn't appear people think that climate change has really impacted their ability to catch fish
#does this mean they think their fishing ability is indepent of climate change?
belief$ability <- factor(belief$ability, levels = c("no", "no_obs_change", "yes_neg", "yes_pos"))

belief$age <- factor(belief$age, levels = c("U30", "30-40", "40-50", "50-60", "60-70", "70+"))

cc_ability_by_age_plot <- ggplot(data = belief, aes (x=ability, fill = age))+
  geom_bar(position = position_dodge(), stat="count")+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))+ ylab("Number of Respondents\n")+
  xlab("Category\n")+
  ggtitle("Climate change has affected ability to catch fish") 

ggsave(plot = cc_ability_by_age_plot, file = paste0("../figures/resp_sum/climate_change_ability_age.png"))

```
#see vulnerability code for q23-32
#see wellbeing code for q33 - 43


#saves data with state responses included
```{r save data}
write.csv(responses, paste0("../data/PNW_decoded_responses_wstate", format(Sys.time(), "%d.%b.%Y"),".csv"))
```
