---
title: "data exploration"
author: "Jocelyn Runnebaum"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
   word_document:
    reference_docx: word-styles-reference.docx

---

```{r setup, echo = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message = FALSE, error=FALSE)
```

```{r, packages}
library(tidyverse)
library(lubridate)
library(knitr)
library(grDevices)
library(extrafont)
library(ggrepel)
library(ggthemes)
library(inauguration)
# windowsFonts(Times=windowsFont("Calibri"))
theme_sleek <- function(base_size = 12, base_family = "Calibri") {
  half_line <- base_size/2
  theme_light(base_size = 12, base_family = "Calibri") +
    theme(
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      axis.ticks.length = unit(half_line / 2.2, "pt"),
      strip.background = element_rect(fill = NA, colour = NA),
      strip.text.x = element_text(colour = "black"),
      strip.text.y = element_text(colour = "black"),
      # axis.text = element_text(colour = "grey30"),
      #axis.title = element_text(colour = "grey30"),
      #legend.title = element_text(colour = "grey30"),#, size = rel(0.9)
      legend.title = element_blank(),
      panel.border = element_rect(fill = NA),#, colour = "grey70", size = 1),
      legend.key.size = unit(0.9, "lines"),
      #legend.text = element_text(size = rel(0.7)),#, colour = "grey30"),
      legend.key = element_rect(colour = NA, fill = NA),
      legend.background = element_rect(colour = NA, fill = NA)#,
      #plot.title = element_text(colour = "grey30"),#, size = rel(1),
      #plot.subtitle = element_text(colour = "grey30")#, size = rel(.85),
     
      
    )
}
theme_set(theme_sleek())
options(scipen = 999) #turns off scientific notation
# Depends on dplyr
tickr <- function(
  data, # dataframe
  var, # column of interest
  to # break point definition 
){
  
  VAR <- enquo(var) # makes VAR a dynamic variable
  
  data %>% 
    distinct(!!VAR) %>%
    ungroup(!!VAR) %>% 
    mutate(labels = ifelse(!!VAR %in% seq(to * round(min(!!VAR) / to), max(!!VAR), to),
                           !!VAR, "")) %>%
    select(breaks = UQ(VAR), labels)
}
```

```{r load data}
responses <- read.csv("../data/decoded_responses_05.Jan.2021.csv")

#add leading zeros back to zip codes in responses dataframe (they drop off when saving from data cleaning)

for(i in 1:length(responses$home_zip)){
  if(as.numeric(responses$home_zip[i]) < 10000){
    responses$home_zip[i] <- paste0("0", responses$home_zip[i])
  }
}

for(i in 1:length(responses$port_zip)){
  if(as.numeric(responses$port_zip[i]) < 10000){
    responses$port_zip[i] <- paste0("0", responses$port_zip[i])
  }
}
```

```{r zipcode data}
#download zipcode data to match states to responses data by zip code (could add county or other identifying information from these data)

ZipCodeSourceFile = "http://download.geonames.org/export/zip/US.zip"
temp <- tempfile()
download.file(ZipCodeSourceFile , temp)
ZipCodes <- read.table(unz(temp, "US.txt"), sep="\t")
unlink(temp)
names(ZipCodes) = c("CountryCode", "port_zip", "PlaceName", 
                    "state", "state_code", "County", "AdminCode2", 
                    "AdminName3", "AdminCode3", "latitude", "longitude", "accuracy")

zips<- ZipCodes %>% 
  select(port_zip, state, state_code) %>% 
  mutate(port_zip = as.character(port_zip))

#add leading zero back
for(i in 1:length(zips$port_zip)){
  if(as.numeric(zips$port_zip[i]) < 10000){
    zips$port_zip[i] <- paste0("0", zips$port_zip[i])
  }
}
```

```{r add state data to responses}

responses <- responses %>% 
  left_join(zips) %>% 
   droplevels() 

#for some reason rows 57 and 67 have unidentifiable zipcodes

#57 = MA
responses[57, 290] = "Massachusetts"
responses[57, 291] = "MA"

#67 = RI
responses[67, 290] = "Rhode Island"
responses[67, 291] = "RI"

belief <- responses %>% 
  select(age, state, state_code, ability, believe, harm_me, harm_future) %>% 
   droplevels() 

```

#Responses by State
```{r responses by State}
#As of 7.27.2020 we only have one response from New Jersey, should we not report data at a level that doesn't have at least 3 observations for the potential of someone accidentally being identified?
responses$state <- factor(responses$state, levels = c("New Jersey","New York", "Connecticut", "Rhode Island", "Massachusetts", "New Hampshire", "Maine"))

responses$state_code <- factor(responses$state_code, levels = c("NJ","NY", "CT", "RI", "MA", "NH", "ME"))

responses_by_state_plot <- ggplot(data = responses, aes(x=state_code))+
  geom_bar(stat = "count")+
  xlab("State\n")+
  ggtitle("Respondents by State") 

ggsave(plot = responses_by_state_plot, file = paste0("../figures/exploration/response_by_state.png"))

```

##Belief in climate change by State
```{r belief in climate change by State}

belief$believe <- factor(belief$believe, levels = c("strongly_disagree", "somewhat_disagree", "neutral", "somewhat_agree", "strongly_agree"))

belief$state <- factor(belief$state, levels = c("New Jersey","New York", "Connecticut", "Rhode Island", "Massachusetts", "New Hampshire", "Maine"))

cc_belief_state_plot <- ggplot(data = belief, aes (x=believe, fill = state))+
  geom_bar(position = position_dodge(), stat="count")+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))+ ylab("Number of Respondents\n")+
  xlab("Category\n")+
  ggtitle("Believe that climate change is occuring") 

ggsave(plot = cc_belief_state_plot, file = paste0("../figures/exploration/climate_change_belief_state.png"))

```

##Believe that climate change will harm them by State
```{r climate change harm me - state}
#Believe that climate change will harm them personally
belief$harm_me <- factor(belief$harm_me, levels = c("strongly_disagree", "somewhat_disagree", "neutral", "somewhat_agree", "strongly_agree"))

belief$state <- factor(belief$state, levels = c("New York", "Connecticut", "Rhode Island", "Massachusetts", "New Hampshire", "Maine"))

cc_harm_me_by_state_plot <- ggplot(data = belief, aes (x=harm_me, fill = state))+
  geom_bar(position = position_dodge(), stat="count")+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))+ ylab("Number of Respondents\n")+
  xlab("Category\n")+
  ggtitle("Believe that climate change will harm them personally") 

ggsave(plot = cc_harm_me_by_state_plot, file = paste0("../figures/exploration/climate_change_harm_me_state.png"))

```

##Believe that climate change will harm future generations by State
```{r climate change harm future - state}
#Believe that climate change will harm future generations
belief$harm_future <- factor(belief$harm_future, levels = c("strongly_disagree", "somewhat_disagree", "neutral", "somewhat_agree", "strongly_agree"))

belief$state <- factor(belief$state, levels = c("New York", "Connecticut", "Rhode Island", "Massachusetts", "New Hampshire", "Maine"))

cc_harm_future_by_state_plot <- ggplot(data = belief, aes (x=harm_future, fill = state))+
  geom_bar(position = position_dodge(), stat="count")+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))+ ylab("Number of Respondents\n")+
  xlab("Category\n")+
  ggtitle("Believe that climate change will harm future generations") 

ggsave(plot = cc_harm_future_by_state_plot, file = paste0("../figures/exploration/climate_change_harm_future_state.png"))

```

#Responses by Age
```{r responses by State}
responses$age <- factor(responses$age, levels = c("U30", "30-40", "40-50", "50-60", "60-70", "70+"))


responses_by_age_plot <- ggplot(data = responses, aes(x=age))+
  geom_bar(position = position_dodge(),stat = "count")
```

##Belief in climate change by age
```{r belief in climate change}
belief$believe <- factor(belief$believe, levels = c("strongly_disagree", "somewhat_disagree", "neutral", "somewhat_agree", "strongly_agree"))
belief$age <- factor(belief$age, levels = c("U30", "30-40", "40-50", "50-60", "60-70", "70+"))


cc_belief_plot <- ggplot(data = belief, aes (x=believe, fill = age))+
  geom_bar(position = position_dodge(), stat="count")+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))+ ylab("Number of Respondents\n")+
  xlab("Category\n")+
  ggtitle("Believe that climate change is occuring") 

ggsave(plot = cc_belief_plot, file = paste0("../figures/climate_change_belief_age.png"))

```

##Believe that climate change will harm them by age
```{r climate change harm me - age}
#Believe that climate change will harm them personally
belief$harm_me <- factor(belief$harm_me, levels = c("strongly_disagree", "somewhat_disagree", "neutral", "somewhat_agree", "strongly_agree"))

belief$age <- factor(belief$age, levels = c("U30", "30-40", "40-50", "50-60", "60-70", "70+"))

cc_harm_me_by_age_plot <- ggplot(data = belief, aes (x=harm_me, fill = age))+
  geom_bar(position = position_dodge(), stat="count")+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))+ ylab("Number of Respondents\n")+
  xlab("Category\n")+
  ggtitle("Believe that climate change will harm them personally") 

ggsave(plot = cc_harm_me_by_age_plot, file = paste0("../figures/exploration/climate_change_harm_me_age.png"))

```

##Believe that climate change will harm future generations by age
```{r climate change harm future - age}
#Believe that climate change will harm future generations
belief$harm_future <- factor(belief$harm_future, levels = c("strongly_disagree", "somewhat_disagree", "neutral", "somewhat_agree", "strongly_agree"))

belief$age <- factor(belief$age, levels = c("U30", "30-40", "40-50", "50-60", "60-70", "70+"))

cc_harm_future_by_age_plot <- ggplot(data = belief, aes (x=harm_future, fill = age))+
  geom_bar(position = position_dodge(), stat="count")+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))+ ylab("Number of Respondents\n")+
  xlab("Category\n")+
  ggtitle("Believe that climate change will harm future generations") 

ggsave(plot = cc_harm_future_by_age_plot, file = paste0("../figures/exploration/climate_change_harm_future_age.png"))

```


##ability to catch fish has been affected by climate change
```{r ability to catch fish}
#Do you think your ability to catch fish has been affected by climate change?
#This is an interesting plot, it doesn't appear people think that climate change has really impacted their ability to catch fish
#does this mean they think their fishing ability is indepent of climate change?
belief$ability <- factor(belief$ability, levels = c("no", "no_obs_change", "yes_neg", "yes_pos"))

belief$age <- factor(belief$age, levels = c("U30", "30-40", "40-50", "50-60", "60-70", "70+"))

cc_ability_by_age_plot <- ggplot(data = belief, aes (x=ability, fill = age))+
  geom_bar(position = position_dodge(), stat="count")+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))+ ylab("Number of Respondents\n")+
  xlab("Category\n")+
  ggtitle("Climate change has affected ability to catch fish") 

ggsave(plot = cc_ability_by_age_plot, file = paste0("../figures/exploration/climate_change_ability_age.png"))

```


```{r}

regions <-  responses %>%
  select(inshore_gom, offshore_gom, gb, sne, lis, mid) %>%  #other_region is also an option
  gather(region, response) %>% 
  group_by(region)

regions$region <- factor(regions$region, levels = c("inshore_gom", "offshore_gom", "gb", "sne", "lis", "mid"))

region_plot <- ggplot()+
  geom_col(data = regions, aes(x=region, y = response), fill = "#749dae")+
  xlab("Species\n")+
  ylab("Count\n")+
  ggtitle("Regions Fished\n")

ggsave(plot = region_plot, file = paste0("../figures/exploration/regions_fished.png"))

```
