---
title: "Species Changes"
author: "Jocelyn"
date: "1/6/2021"
output: html_document
---

```{r setup, echo = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message = FALSE, error=FALSE)
```

```{r, packages}
library(tidyverse)
library(lubridate)
library(knitr)
library(grDevices)
library(extrafont)
library(ggrepel)
library(ggthemes)
library(gridExtra)
library(likert)
# library(devtools)
# devtools::install_github("ciannabp/inauguration")
library(inauguration)

# windowsFonts(Times=windowsFont("Calibri"))
theme_sleek <- function(base_size = 12, base_family = "Calibri") {
  half_line <- base_size/2
  theme_light(base_size = 12, base_family = "Calibri") +
    theme(
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      axis.ticks.length = unit(half_line / 2.2, "pt"),
      strip.background = element_rect(fill = NA, colour = NA),
      strip.text.x = element_text(colour = "black"),
      strip.text.y = element_text(colour = "black"),
      # axis.text = element_text(colour = "grey30"),
      #axis.title = element_text(colour = "grey30"),
      #legend.title = element_text(colour = "grey30"),#, size = rel(0.9)
      legend.title = element_blank(),
      panel.border = element_rect(fill = NA),#, colour = "grey70", size = 1),
      legend.key.size = unit(0.9, "lines"),
      #legend.text = element_text(size = rel(0.7)),#, colour = "grey30"),
      legend.key = element_rect(colour = NA, fill = NA),
      legend.background = element_rect(colour = NA, fill = NA)#,
      #plot.title = element_text(colour = "grey30"),#, size = rel(1),
      #plot.subtitle = element_text(colour = "grey30")#, size = rel(.85),
     
      
    )
}
theme_set(theme_sleek())
options(scipen = 999) #turns off scientific notation
# Depends on dplyr
tickr <- function(
  data, # dataframe
  var, # column of interest
  to # break point definition 
){
  
  VAR <- enquo(var) # makes VAR a dynamic variable
  
  data %>% 
    distinct(!!VAR) %>%
    ungroup(!!VAR) %>% 
    mutate(labels = ifelse(!!VAR %in% seq(to * round(min(!!VAR) / to), max(!!VAR), to),
                           !!VAR, "")) %>%
    select(breaks = UQ(VAR), labels)
}
```

#looking at changes in species 
```{r load data}
responses <- read.csv("../data/decoded_responses_05.Jan.2021.csv")
```

         

```{r question 23}
#q23 plotting using likert package
#What, if any, effect do you believe ocean warming is having on these fisheries?

q23<-responses %>%
  select(lob_warming, herr_warming, men_warming, sclp_warming, cod_warming, hadk_warming, polk_warming, halbt_warming, wht_hake_warming, dabs_warming, gry_sole_warming, windowpane_warming, yellowtail_warming, wnt_flnder_warming, redfish_warming, monkfish_warming, skates_warming, whiting_warming, red_hake_warming, red_crab_warming, clams_warming, squid_warming, dogfish_warming, summer_flnder_warming, scup_warming, black_seabass_warming)

#add statements as column headers
q23_statements<-c(
  "Lobster",
  "Atlantic Herring",
  "Menhaden",
  "Scallop",
  "Atlantic Cod",
  "Haddock",
  "Pollock",
  "Halibut",
  "White Hake",
  "Dabs",
  "Gray Sole",
  "Windowpane Flounder",
  "Yellowtail Flounder",
  "Winter Flounder",
  "Acadian Redfish",
  "Monkfish",
  "Skates",
  "Whiting",
  "Red Hake",
  "Red Crab",
  "Clams",
  "Squid",
  "Dogfish",
  "Summer Flounder",
  "Scup",
  "Black Sea Bass")

colnames(q23)<-q23_statements

#setting the levels
#levels need to be set since each possible answer does not appear in each column

mylevels <- c('strong_neg' ,'slight_neg', 'no_effect', 'slight_pos', 'strong_pos')
for(i in seq_along(q23)) {
  q23[,i] <- factor(q23[,i], levels=mylevels)
}

species_change <- likert(q23)
species_title <- "Percieved effect of ocean warming on fisheries"
summary(species_change)
species_change_plot <- plot(species_change, colors = rev(inauguration("inauguration_2021", 5)), ordered=FALSE, group.order=names(q23)) + ggtitle(species_title)
                          
ggsave(plot = species_change_plot, file = paste0("../figures/species_change/species_change.png"))
```


```{r question 24}
#q24 plotting using likert package
#Please indicate your level of confidence in your responses to the previous question.
q24<-responses %>%
  select(lob_warming_conf, herr_warming_conf, men_warming_conf, sclp_warming_conf, cod_warming_conf, hadk_warming_conf, polk_warming_conf, halbt_warming_conf, wht_hake_warming_conf, dabs_warming_conf, gry_sole_warming_conf, windowpane_warming_conf, yellowtail_warming_conf, wnt_flnder_warming_conf, redfish_warming_conf, monkfish_warming_conf, skates_warming_conf, whiting_warming_conf, red_hake_warming_conf, red_crab_warming_conf, clams_warming_conf, squid_warming_conf, dogfish_warming_conf, summer_flnder_warming_conf, scup_warming_conf, black_seabass_warming_conf)

#add statements as column headers
q24_statements<-c(
  "Lobster",
  "Atlantic Herring",
  "Menhaden",
  "Scallop",
  "Atlantic Cod",
  "Haddock",
  "Pollock",
  "Halibut",
  "White Hake",
  "Dabs",
  "Gray Sole",
  "Windowpade Flounder",
  "Yellowtail Flounder",
  "Winter Flounder",
  "Acadian Redfish",
  "Monkfish",
  "Skates",
  "Whiting",
  "Red Hake",
  "Red Crab",
  "Clams",
  "Squid",
  "Dogfish",
  "Summer Flounder",
  "Scup",
  "Black Sea Bass")

colnames(q24)<-q24_statements

#setting the levels
#levels need to be set since each possible answer does not appear in each column

mylevels <- c('low' ,'med', 'high')
for(i in seq_along(q24)) {
  q24[,i] <- factor(q24[,i], levels=mylevels)
}


species_change_conf <- likert(q24)

species_change_conf_title <- "Confidence in percieved effect of ocean warming on fisheries"
summary(species_change_conf)
species_change_conf_plot <- plot(species_change_conf, ordered=FALSE, group.order=names(q24), colors = rev(inauguration("inauguration_2021", 3))) + ggtitle(species_change_conf_title)
                          
ggsave(plot = species_change_conf_plot, file = paste0("../figures/species_change/species_change_conf.png"))
```

#look at perceptions by State
```{r question 23, Maine}
maine <- responses %>% 
  filter(state == "Maine")

#q23 plotting using likert package
#What, if any, effect do you believe ocean warming is having on these fisheries?
q23_maine <- maine %>%
  select(lob_warming, herr_warming, men_warming, sclp_warming, cod_warming, hadk_warming, polk_warming, halbt_warming, wht_hake_warming, dabs_warming, gry_sole_warming, windowpane_warming, yellowtail_warming, wnt_flnder_warming, redfish_warming, monkfish_warming, skates_warming, whiting_warming, red_hake_warming, red_crab_warming, clams_warming, squid_warming, dogfish_warming, summer_flnder_warming, scup_warming, black_seabass_warming)

#add statements as column headers

colnames(q23_maine)<-q23_statements

#setting the levels
#levels need to be set since each possible answer does not appear in each column

mylevels <- c('strong_neg' ,'slight_neg', 'no_effect', 'slight_pos', 'strong_pos')
for(i in seq_along(q23_maine)) {
  q23_maine[,i] <- factor(q23_maine[,i], levels=mylevels)
}

species_change_maine <- likert(q23_maine)
species_title_maine <- "Maine perceptions"
summary(species_change_maine)
species_change_maine_plot <- plot(species_change_maine, ordered=FALSE, group.order=names(q23), colors = rev(inauguration("inauguration_2021", 5))) + ggtitle(species_title_maine)
                          
ggsave(plot = species_change_maine_plot, file = paste0("../figures/species_change/species_change_maine.png"))
```

```{r question 23, New Hampshire}
nh <- responses %>% 
  filter(state == "New Hampshire")

#q23 plotting using likert package
#What, if any, effect do you believe ocean warming is having on these fisheries?
q23_nh <- nh %>%
  select(lob_warming, herr_warming, men_warming, sclp_warming, cod_warming, hadk_warming, polk_warming, halbt_warming, wht_hake_warming, dabs_warming, gry_sole_warming, windowpane_warming, yellowtail_warming, wnt_flnder_warming, redfish_warming, monkfish_warming, skates_warming, whiting_warming, red_hake_warming, red_crab_warming, clams_warming, squid_warming, dogfish_warming, summer_flnder_warming, scup_warming, black_seabass_warming)

#add statements as column headers

colnames(q23_nh)<-q23_statements

#setting the levels
#levels need to be set since each possible answer does not appear in each column

mylevels <- c('strong_neg' ,'slight_neg', 'no_effect', 'slight_pos', 'strong_pos')
for(i in seq_along(q23_nh)) {
  q23_nh[,i] <- factor(q23_nh[,i], levels=mylevels)
}

species_change_nh <- likert(q23_nh)
species_title_nh <- "New Hampshire perceptions"
summary(species_change_nh)
species_change_nh_plot <- plot(species_change_nh, ordered=FALSE, group.order=names(q23), colors = rev(inauguration("inauguration_2021", 5))) + ggtitle(species_title_nh)
                          
ggsave(plot = species_change_nh_plot, file = paste0("../figures/species_change/species_change_nh.png"))
```

```{r question 23, Massachusetts}
mass <- responses %>% 
  filter(state == "Massachusetts")

#q23 plotting using likert package
#What, if any, effect do you believe ocean warming is having on these fisheries?
q23_mass <- mass %>%
  select(lob_warming, herr_warming, men_warming, sclp_warming, cod_warming, hadk_warming, polk_warming, halbt_warming, wht_hake_warming, dabs_warming, gry_sole_warming, windowpane_warming, yellowtail_warming, wnt_flnder_warming, redfish_warming, monkfish_warming, skates_warming, whiting_warming, red_hake_warming, red_crab_warming, clams_warming, squid_warming, dogfish_warming, summer_flnder_warming, scup_warming, black_seabass_warming)

#add statements as column headers

colnames(q23_mass)<-q23_statements

#setting the levels
#levels need to be set since each possible answer does not appear in each column

mylevels <- c('strong_neg' ,'slight_neg', 'no_effect', 'slight_pos', 'strong_pos')
for(i in seq_along(q23_mass)) {
  q23_mass[,i] <- factor(q23_mass[,i], levels=mylevels)
}

species_change_mass <- likert(q23_mass)
species_title_mass <- "Massachusetts perceptions"
summary(species_change_mass)
species_change_mass_plot <- plot(species_change_mass, ordered=FALSE, group.order=names(q23), colors = rev(inauguration("inauguration_2021", 5))) + ggtitle(species_title_mass)
                          
ggsave(plot = species_change_mass_plot, file = paste0("../figures/species_change/species_change_mass.png"))
```

```{r question 23, Rhode Island}
ri <- responses %>% 
  filter(state == "Rhode Island")

#q23 plotting using likert package
#What, if any, effect do you believe ocean warming is having on these fisheries?
q23_ri <- ri %>%
  select(lob_warming, herr_warming, men_warming, sclp_warming, cod_warming, hadk_warming, polk_warming, halbt_warming, wht_hake_warming, dabs_warming, gry_sole_warming, windowpane_warming, yellowtail_warming, wnt_flnder_warming, redfish_warming, monkfish_warming, skates_warming, whiting_warming, red_hake_warming, red_crab_warming, clams_warming, squid_warming, dogfish_warming, summer_flnder_warming, scup_warming, black_seabass_warming)

#add statements as column headers

colnames(q23_ri)<-q23_statements

#setting the levels
#levels need to be set since each possible answer does not appear in each column

mylevels <- c('strong_neg' ,'slight_neg', 'no_effect', 'slight_pos', 'strong_pos')
for(i in seq_along(q23_ri)) {
  q23_ri[,i] <- factor(q23_ri[,i], levels=mylevels)
}

species_change_ri <- likert(q23_ri)
species_title_ri <- "Rhode Island perceptions"
summary(species_change_ri)
species_change_ri_plot <- plot(species_change_ri, ordered=FALSE, group.order=names(q23), colors = rev(inauguration("inauguration_2021", 5))) + ggtitle(species_title_ri)
                          
ggsave(plot = species_change_ri_plot, file = paste0("../figures/species_change/species_change_ri.png"))
```

```{r question 23, Connecticut}
ct <- responses %>% 
  filter(state == "Connecticut")

#q23 plotting using likert package
#What, if any, effect do you believe ocean warming is having on these fisheries?
q23_ct <- ct %>%
  select(lob_warming, herr_warming, men_warming, sclp_warming, cod_warming, hadk_warming, polk_warming, halbt_warming, wht_hake_warming, dabs_warming, gry_sole_warming, windowpane_warming, yellowtail_warming, wnt_flnder_warming, redfish_warming, monkfish_warming, skates_warming, whiting_warming, red_hake_warming, red_crab_warming, clams_warming, squid_warming, dogfish_warming, summer_flnder_warming, scup_warming, black_seabass_warming)

#add statements as column headers

colnames(q23_ct)<-q23_statements

#setting the levels
#levels need to be set since each possible answer does not appear in each column

mylevels <- c('strong_neg' ,'slight_neg', 'no_effect', 'slight_pos', 'strong_pos')
for(i in seq_along(q23_ct)) {
  q23_ct[,i] <- factor(q23_ct[,i], levels=mylevels)
}

species_change_ct <- likert(q23_ct)
species_title_ct <- "Connecticut perceptions"
summary(species_change_ct)
species_change_ct_plot <- plot(species_change_ct, ordered=FALSE, group.order=names(q23),colors = rev(inauguration("inauguration_2021", 5))) + ggtitle(species_title_ct)
                          
ggsave(plot = species_change_ct_plot, file = paste0("../figures/species_change/species_change_ct.png"))
```

#look at perceptions by fishing region
responses <- responses %>% 
  rename(inshore_gom = q3r1, #inshore gulf of maine (defined by the 50 fathom curve)
         offshore_gom = q3r2, #offshore gulf of maine (defined by the 50 fathom curve)
         gb = q3r3, #georges bank
         sne = q3r4, #southern new england
         lis = q3r5, #long island sound
         mid = q3r6, #mid atlantic
         other_region = q3r7, 
         other_region_descrip = q3r7oe) #defined by participant
```{r question 23, inshore GOM}
inshore_gom <- responses %>% 
  filter(inshore_gom == 1)

#q23 plotting using likert package
#What, if any, effect do you believe ocean warming is having on these fisheries?
q23_inshore_gom <- inshore_gom %>%
   select(lob_warming, windowpane_warming, whiting_warming, summer_flnder_warming ) #look at harvester/scientist disagreement on impact
#look at the impact of climate change on all species included in the survey
# lob_warming, herr_warming, men_warming, sclp_warming, cod_warming, hadk_warming, polk_warming, halbt_warming, wht_hake_warming, dabs_warming, gry_sole_warming, windowpane_warming, yellowtail_warming, wnt_flnder_warming, redfish_warming, monkfish_warming, skates_warming, whiting_warming, red_hake_warming, red_crab_warming, clams_warming, squid_warming, dogfish_warming, summer_flnder_warming, scup_warming, black_seabass_warming

#add statements as column headers

q23_disagreement_statements<-c(
  "Lobster",
  "Windowpade Flounder",
  "Whiting",
  "Summer Flounder")

q23

#add statements as column headers

colnames(q23_inshore_gom) <- q23_disagreement_statements #change this to q23_statements to run for all species assessed. 
  # q23_statements

#setting the levels
#levels need to be set since each possible answer does not appear in each column

mylevels <- c('strong_neg' ,'slight_neg', 'no_effect', 'slight_pos', 'strong_pos')
for(i in seq_along(q23_inshore_gom)) {
  q23_inshore_gom[,i] <- factor(q23_inshore_gom[,i], levels=mylevels)
}

species_change_inshore_GOM <- likert(q23_inshore_gom)
species_title_inshore_GOM <- "Inshore Gulf of Maine Perceptions"
summary(species_change_inshore_GOM)
species_change_inshore_gom_plot <- plot(species_change_inshore_GOM, ordered=FALSE, group.order=names(q23_inshore_gom), colors = rev(inauguration("inauguration_2021", 5))) + ggtitle(species_title_inshore_GOM)
                          
ggsave(plot = species_change_inshore_gom_plot, file = paste0("../figures/species_change/species_change_inshore_gom_disagreement.png"))
# "../figures/species_change/species_change_inshore_gom.png"
```

```{r question 23, offshore GOM}
offshore_gom <- responses %>% 
  filter(offshore_gom == 1)

#q23 plotting using likert package
#What, if any, effect do you believe ocean warming is having on these fisheries?
q23_offshore_gom <- offshore_gom %>%
  select(lob_warming, windowpane_warming, whiting_warming, summer_flnder_warming ) #look at harvester/scientist disagreement on impact
#look at the impact of climate change on all species included in the survey
# lob_warming, herr_warming, men_warming, sclp_warming, cod_warming, hadk_warming, polk_warming, halbt_warming, wht_hake_warming, dabs_warming, gry_sole_warming, windowpane_warming, yellowtail_warming, wnt_flnder_warming, redfish_warming, monkfish_warming, skates_warming, whiting_warming, red_hake_warming, red_crab_warming, clams_warming, squid_warming, dogfish_warming, summer_flnder_warming, scup_warming, black_seabass_warming

#add statements as column headers

q23_disagreement_statements<-c(
  "Lobster",
  "Windowpade Flounder",
  "Whiting",
  "Summer Flounder")

colnames(q23_offshore_gom) <- q23_disagreement_statements
  # q23_statements

#setting the levels
#levels need to be set since each possible answer does not appear in each column

mylevels <- c('strong_neg' ,'slight_neg', 'no_effect', 'slight_pos', 'strong_pos')
for(i in seq_along(q23_offshore_gom)) {
  q23_offshore_gom[,i] <- factor(q23_offshore_gom[,i], levels=mylevels)
}

species_change_offshore_gom <- likert(q23_offshore_gom)
species_title_offshore_gom <- "Offshore Gulf of Maine Perceptions"
summary(species_change_offshore_gom)
species_change_offshore_gom_plot <- plot(species_change_offshore_gom, ordered=FALSE, group.order=names(q23_offshore_gom), colors = rev(inauguration("inauguration_2021", 5))) + ggtitle(species_title_offshore_gom)
                          
ggsave(plot = species_change_offshore_gom_plot, file = paste0("../figures/species_change/species_change_offshore_gom_disagreement.png"))
# "../figures/species_change/species_change_offshore_gom.png"
```

```{r question 23, Georges Bank}
gb <- responses %>% 
  filter(gb == 1)

#q23 plotting using likert package
#What, if any, effect do you believe ocean warming is having on these fisheries?
q23_gb <- gb %>%
   select(lob_warming, windowpane_warming, whiting_warming, summer_flnder_warming ) #look at harvester/scientist disagreement on impact
#look at the impact of climate change on all species included in the survey
# lob_warming, herr_warming, men_warming, sclp_warming, cod_warming, hadk_warming, polk_warming, halbt_warming, wht_hake_warming, dabs_warming, gry_sole_warming, windowpane_warming, yellowtail_warming, wnt_flnder_warming, redfish_warming, monkfish_warming, skates_warming, whiting_warming, red_hake_warming, red_crab_warming, clams_warming, squid_warming, dogfish_warming, summer_flnder_warming, scup_warming, black_seabass_warming

#add statements as column headers

q23_disagreement_statements<-c(
  "Lobster",
  "Windowpade Flounder",
  "Whiting",
  "Summer Flounder")

colnames(q23_gb) <- q23_disagreement_statements
  # q23_statements

#setting the levels
#levels need to be set since each possible answer does not appear in each column

mylevels <- c('strong_neg' ,'slight_neg', 'no_effect', 'slight_pos', 'strong_pos')
for(i in seq_along(q23_gb)) {
  q23_gb[,i] <- factor(q23_gb[,i], levels=mylevels)
}

species_change_gb <- likert(q23_gb)
species_title_gb <- "Georges Bank Perceptions"
summary(species_change_gb)
species_change_gb_plot <- plot(species_change_gb, ordered=FALSE, group.order=names(q23_gb), colors = rev(inauguration("inauguration_2021", 5))) + ggtitle(species_title_gb)
                          
ggsave(plot = species_change_gb_plot, file = paste0("../figures/species_change/species_change_gb_disagreement.png"))
"../figures/species_change/species_change_gb.png"
```

```{r question 23, Southern New England}
sne <- responses %>% 
  filter(sne == 1)

#q23 plotting using likert package
#What, if any, effect do you believe ocean warming is having on these fisheries?
q23_sne <- sne %>%
  select(lob_warming, windowpane_warming, whiting_warming, summer_flnder_warming ) #look at harvester/scientist disagreement on impact
#look at the impact of climate change on all species included in the survey
# lob_warming, herr_warming, men_warming, sclp_warming, cod_warming, hadk_warming, polk_warming, halbt_warming, wht_hake_warming, dabs_warming, gry_sole_warming, windowpane_warming, yellowtail_warming, wnt_flnder_warming, redfish_warming, monkfish_warming, skates_warming, whiting_warming, red_hake_warming, red_crab_warming, clams_warming, squid_warming, dogfish_warming, summer_flnder_warming, scup_warming, black_seabass_warming

#add statements as column headers
q23_disagreement_statements<-c(
  "Lobster",
  "Windowpade Flounder",
  "Whiting",
  "Summer Flounder")

colnames(q23_sne) <- q23_disagreement_statements
  # q23_statements

#setting the levels
#levels need to be set since each possible answer does not appear in each column

mylevels <- c('strong_neg' ,'slight_neg', 'no_effect', 'slight_pos', 'strong_pos')
for(i in seq_along(q23_sne)) {
  q23_sne[,i] <- factor(q23_sne[,i], levels=mylevels)
}

species_change_sne <- likert(q23_sne)
species_title_sne <- "Southern New England Perceptions"
summary(species_change_sne)
species_change_sne_plot <- plot(species_change_sne, ordered=FALSE, group.order=names(q23_sne), colors = rev(inauguration("inauguration_2021", 5))) + ggtitle(species_title_sne)
                          
ggsave(plot = species_change_sne_plot, file = paste0("../figures/species_change/species_change_sne_disagreement.png"))
# "../figures/species_change/species_change_sne.png"
```

```{r question 23, Long Island Sound}
lis <- responses %>% 
  filter(lis == 1)

#q23 plotting using likert package
#What, if any, effect do you believe ocean warming is having on these fisheries?
q23_lis <- lis %>%
 select(lob_warming, windowpane_warming, whiting_warming, summer_flnder_warming ) #look at harvester/scientist disagreement on impact
#look at the impact of climate change on all species included in the survey
# lob_warming, herr_warming, men_warming, sclp_warming, cod_warming, hadk_warming, polk_warming, halbt_warming, wht_hake_warming, dabs_warming, gry_sole_warming, windowpane_warming, yellowtail_warming, wnt_flnder_warming, redfish_warming, monkfish_warming, skates_warming, whiting_warming, red_hake_warming, red_crab_warming, clams_warming, squid_warming, dogfish_warming, summer_flnder_warming, scup_warming, black_seabass_warming

#add statements as column headers

q23_disagreement_statements<-c(
  "Lobster",
  "Windowpade Flounder",
  "Whiting",
  "Summer Flounder")

colnames(q23_lis) <- q23_disagreement_statements
  # q23_statements

#setting the levels
#levels need to be set since each possible answer does not appear in each column

mylevels <- c('strong_neg' ,'slight_neg', 'no_effect', 'slight_pos', 'strong_pos')
for(i in seq_along(q23_lis)) {
  q23_lis[,i] <- factor(q23_lis[,i], levels=mylevels)
}

species_change_lis <- likert(q23_lis)
species_title_lis <- "Long Island Sound Perceptions"
summary(species_change_lis)
species_change_lis_plot <- plot(species_change_lis, ordered=FALSE, group.order=names(q23_lis), colors = rev(inauguration("inauguration_2021", 5))) + ggtitle(species_title_lis)
                          
ggsave(plot = species_change_lis_plot, file = paste0("../figures/species_change/species_change_lis_disgreement.png"))
# 
```

```{r question 23, Mid-Atlantic}
mid <- responses %>% 
  filter(mid == 1)

#q23 plotting using likert package
#What, if any, effect do you believe ocean warming is having on these fisheries?
q23_mid <- mid %>%
  select(lob_warming, windowpane_warming, whiting_warming, summer_flnder_warming ) #look at harvester/scientist disagreement on impact
#look at the impact of climate change on all species included in the survey
# lob_warming, herr_warming, men_warming, sclp_warming, cod_warming, hadk_warming, polk_warming, halbt_warming, wht_hake_warming, dabs_warming, gry_sole_warming, windowpane_warming, yellowtail_warming, wnt_flnder_warming, redfish_warming, monkfish_warming, skates_warming, whiting_warming, red_hake_warming, red_crab_warming, clams_warming, squid_warming, dogfish_warming, summer_flnder_warming, scup_warming, black_seabass_warming

#add statements as column headers
q23_disagreement_statements<-c(
  "Lobster",
  "Windowpade Flounder",
  "Whiting",
  "Summer Flounder")

colnames(q23_mid) <- q23_disagreement_statements
  # q23_statements

#setting the levels
#levels need to be set since each possible answer does not appear in each column

mylevels <- c('strong_neg' ,'slight_neg', 'no_effect', 'slight_pos', 'strong_pos')
for(i in seq_along(q23_mid)) {
  q23_mid[,i] <- factor(q23_mid[,i], levels=mylevels)
}

species_change_mid <- likert(q23_mid)
species_title_mid <- "Mid-Atlantic Perceptions"
summary(species_change_mid)
species_change_mid_plot <- plot(species_change_mid, ordered=FALSE, group.order=names(q23_mid), colors = rev(inauguration("inauguration_2021", 5))) + ggtitle(species_title_mid)
                          
ggsave(plot = species_change_mid_plot, file = paste0("../figures/species_change/species_change_mid_disagreement.png"))
# "../figures/species_change/species_change_mid.png"
```

