---
title: "Relative concerns for ac"
author: "Laura"
date: "1/19/2022"
output: html_document
---


#concerns table
#responses from questions 33-36 about issues that may affect fishing success and wellbeing
#grouped into categories following the survey: marine environment, fishing operations, community and infrastructure, and personal
```{r}

worries<-responses %>%
  select(warming_concern, oa_concern, storm_concern, sea_level_concern, weather_concern, h20_quality_concern, algal_concern, habitat_concern, pop_size_concern, bycatch_concern, value_concern, costs_concern, sa_concern, distance_concern, reg_concern, graying_fleet_concern, fish_comm_concern, resid_concern, infrastructure_concern, phsy_health_concern, mental_health_concern, safe_at_sea_concern, family_concern, think_warming, think_oa, think_storm, think_sea_level, think_weather, think_h20_quality, think_algal, think_habitat, think_pop_size, think_bycatch, think_value, think_costs, think_sa, think_distance, think_reg, think_graying_fleet, think_fish_comm, think_resid, think_infrastructure, think_phsy_health, think_mental_health, think_safe_at_sea, think_family)
```

#function for numberic worry scale
```{r}
worry_calc <- function(x){
if(is.na(x)){
  return("NA")
} else if(x == "none"){
  return(1)
} else if(x == "somewhat"){
  return(2)
} else if(x == "very"){
  return(3)
}
}

freq_calc <- function(x){
if(is.na(x)){
  return("NA")
} else if(x == "never"){
  return(1)
} else if(x == "occasionally"){
  return(2)
} else if(x == "frequently"){
  return(3)
}
}
```

#giving numberic values to responses
```{r}
worries <- worries %>% 
  mutate(warming_concern = sapply(X=worries$warming_concern, FUN=worry_calc),
         oa_concern = sapply(X=worries$oa_concern, FUN=worry_calc),
         storm_concern = sapply(X=worries$storm_concern, FUN=worry_calc),
         sea_level_concern = sapply(X=worries$sea_level_concern, FUN=worry_calc),
         weather_concern = sapply(X=worries$weather_concern, FUN=worry_calc),
         h20_quality_concern = sapply(X=worries$h20_quality_concern, FUN=worry_calc),
         algal_concern = sapply(X=worries$algal_concern, FUN=worry_calc),
         habitat_concern = sapply(X=worries$habitat_concern, FUN=worry_calc),
         pop_size_concern = sapply(X=worries$pop_size_concern, FUN=worry_calc),
         bycatch_concern = sapply(X=worries$bycatch_concern, FUN=worry_calc),
         value_concern = sapply(X=worries$value_concern, FUN=worry_calc),
         costs_concern = sapply(X=worries$costs_concern, FUN=worry_calc),
         sa_concern = sapply(X=worries$sa_concern, FUN=worry_calc),
         distance_concern = sapply(X=worries$distance_concern, FUN=worry_calc),
         reg_concern = sapply(X=worries$reg_concern, FUN=worry_calc),
         graying_fleet_concern = sapply(X=worries$graying_fleet_concern, FUN=worry_calc),
         fish_comm_concern = sapply(X=worries$fish_comm_concern, FUN=worry_calc),
         resid_concern = sapply(X=worries$resid_concern, FUN=worry_calc),
         infrastructure_concern = sapply(X=worries$infrastructure_concern, FUN=worry_calc),
         phsy_health_concern = sapply(X=worries$phsy_health_concern, FUN=worry_calc),
         mental_health_concern = sapply(X=worries$mental_health_concern, FUN=worry_calc),
         safe_at_sea_concern = sapply(X=worries$safe_at_sea_concern, FUN=worry_calc),
         family_concern = sapply(X=worries$family_concern, FUN=worry_calc),
         think_warming = sapply(X=think_warming, FUN = freq_calc),
         think_oa = sapply(X=think_oa, FUN = freq_calc),
         think_storm = sapply(X=think_storm, FUN = freq_calc),
         think_sea_level = sapply(X=think_sea_level, FUN = freq_calc),
         think_weather = sapply(X=think_weather, FUN = freq_calc),
         think_h20_quality = sapply(X=think_h20_quality, FUN = freq_calc),
         think_algal = sapply(X=think_algal, FUN = freq_calc),
         think_habitat = sapply(X=think_habitat, FUN = freq_calc), 
         think_pop_size = sapply(X=think_pop_size, FUN = freq_calc), 
         think_bycatch = sapply(X=think_bycatch, FUN = freq_calc), 
         think_value = sapply(X=think_value, FUN = freq_calc), 
         think_costs = sapply(X=think_costs, FUN = freq_calc), 
         think_sa = sapply(X=think_sa, FUN = freq_calc), 
         think_distance = sapply(X=think_distance, FUN = freq_calc), 
         think_reg = sapply(X=think_reg, FUN = freq_calc), 
         think_graying_fleet = sapply(X=think_graying_fleet, FUN = freq_calc), 
         think_fish_comm = sapply(X=think_fish_comm, FUN = freq_calc), 
         think_resid = sapply(X=think_resid, FUN = freq_calc), 
         think_infrastructure = sapply(X=think_infrastructure, FUN = freq_calc), 
         think_phsy_health = sapply(X=think_phsy_health, FUN = freq_calc), 
         think_mental_health = sapply(X=think_mental_health, FUN = freq_calc), 
         think_safe_at_sea = sapply(X=think_safe_at_sea, FUN = freq_calc), 
         think_family = sapply(X=think_family, FUN = freq_calc))
```

#group by category
```{r}
me<-worries %>%
  select(warming_concern, oa_concern, storm_concern, sea_level_concern, weather_concern, h20_quality_concern, algal_concern, habitat_concern, think_warming, think_oa, think_storm, think_sea_level, think_weather, think_h20_quality, think_algal, think_habitat)

fo<-worries %>%
  select(pop_size_concern, bycatch_concern, value_concern, costs_concern, sa_concern, distance_concern, reg_concern, think_pop_size, think_bycatch, think_value, think_costs, think_sa, think_distance, think_reg)

comin <- worries %>%
  select(graying_fleet_concern, fish_comm_concern, resid_concern, infrastructure_concern,think_graying_fleet, think_fish_comm, think_resid, think_infrastructure)

per<- worries %>%
  select(phsy_health_concern, mental_health_concern, safe_at_sea_concern, family_concern, think_phsy_health, think_mental_health, think_safe_at_sea, think_family)
```
#score for each category
#average of responses for the questions in each category
#weighted by frequency of thought
```{r}
#marine environment
me$marine_env = as.numeric(NA, length(nrow(me)))

me<-me %>%
  mutate(marine_env = (warming_concern*think_warming + 
        oa_concern*think_oa + 
        storm_concern*think_storm +
        sea_level_concern*think_sea_level +
        weather_concern*think_weather +
        h20_quality_concern*think_h20_quality +
        algal_concern*think_algal +
        habitat_concern*think_habitat)/8)

worries$marine_env = me$marine_env

#fishing operations
fo$fishing_ops = as.numeric(NA, length(nrow(fo)))

fo<-fo %>%
  mutate(fishing_ops = 
           (pop_size_concern*think_pop_size +
           bycatch_concern*think_bycatch +
           value_concern*think_value +
           costs_concern*think_costs +
           sa_concern*think_sa +
           distance_concern*think_distance +
           reg_concern*think_reg)/7)

worries$fishing_ops = fo$fishing_ops

#community and infrastructure
comin$community_infras = as.numeric(NA, length(nrow(comin)))

comin <- comin %>%
  mutate(community_infras = 
           (graying_fleet_concern*think_graying_fleet +
           fish_comm_concern*think_fish_comm +
           resid_concern*think_resid +
           infrastructure_concern*think_infrastructure)/4)

worries$community_infras = comin$community_infras

#personal
per$personal = as.numeric(NA, length(nrow(per)))

per<-per %>%
  mutate(personal =
           (phsy_health_concern*think_phsy_health +
           mental_health_concern*think_mental_health +
           safe_at_sea_concern*think_safe_at_sea +
           family_concern*think_family)/4)

worries$personal = per$personal
```