---
title: "social indicator and ac"
author: "Laura"
date: "11/12/2021"
output: html_document
---

```{r setup, echo = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message = FALSE, error=FALSE)
```

```{r, packages}
library(tidyverse)
library(lubridate)
library(knitr)
library(grDevices)
library(extrafont)
library(ggrepel)
library(ggthemes)
library(ggplot2)
library(grid)
library(RColorBrewer)
library(ggpubr)
library(PNWColors)
library(GGally)
library(factoextra)
library(psych)
library(NbClust)
# windowsFonts(Times=windowsFont("Calibri"))
theme_sleek <- function(base_size = 12, base_family = "Calibri") {
  half_line <- base_size/2
  theme_light(base_size = 12, base_family = "Calibri") +
    theme(
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      axis.ticks.length = unit(half_line / 2.2, "pt"),
      strip.background = element_rect(fill = NA, colour = NA),
      strip.text.x = element_text(colour = "black"),
      strip.text.y = element_text(colour = "black"),
      # axis.text = element_text(colour = "grey30"),
      #axis.title = element_text(colour = "grey30"),
      #legend.title = element_text(colour = "grey30"),#, size = rel(0.9)
      legend.title = element_blank(),
      panel.border = element_rect(fill = NA),#, colour = "grey70", size = 1),
      legend.key.size = unit(0.9, "lines"),
      #legend.text = element_text(size = rel(0.7)),#, colour = "grey30"),
      legend.key = element_rect(colour = NA, fill = NA),
      legend.background = element_rect(colour = NA, fill = NA)#,
      #plot.title = element_text(colour = "grey30"),#, size = rel(1),
      #plot.subtitle = element_text(colour = "grey30")#, size = rel(.85),
     
      
    )
}
theme_set(theme_sleek())
options(scipen = 999) #turns off scientific notation
# Depends on dplyr
tickr <- function(
  data, # dataframe
  var, # column of interest
  to # break point definition 
){
  
  VAR <- enquo(var) # makes VAR a dynamic variable
  
  data %>% 
    distinct(!!VAR) %>%
    ungroup(!!VAR) %>% 
    mutate(labels = ifelse(!!VAR %in% seq(to * round(min(!!VAR) / to), max(!!VAR), to),
                           !!VAR, "")) %>%
    select(breaks = UQ(VAR), labels)
}
```

Load NOAA SVI data
```{r}
social <- read.csv("../data/socialIndicatorData.csv")
```

Get just US portion of data since no NOAA stuff for BC
```{r}
us_ac <- ac %>%
  filter(region == "wc" | region == "ne")

ne_ports<- ne %>%
  dplyr::select(ID, homeport)

wc_port <- responses %>%
  dplyr::select(ID, homeport)

ports<- full_join(ne_ports, wc_port)

us_ac<- us_ac %>%
  left_join(ports)

#add on the zips and lat long from other dataframe

llzips<-domainsmod %>%
  dplyr::select(ID, state_code, home_zip, port_zip, latitude, longitude)

us_ac<- us_ac %>%
  left_join(llzips)
```

merging NOAA community SVI data
```{r cleaned port names}

#selecting specific data from noaa stuff
svi<-social %>%
  dplyr::select(Year, Community.Name, State, Commercial.Engagement, Latitude, Longitude, Commercial.Reliance, Commercial.Engagement.Categorical.Ranking, Commercial.Reliance.Categorical.Ranking)  %>%   
  filter(Year == "2018")%>%
  rename(state_code = State,
         homeport = Community.Name)

#get rid of space around state names 
svi$state_code <- trimws(svi$state_code)

trial <- us_ac %>%
  left_join(svi, by = c("homeport", "state_code"))

trial<-domainsmod%>%
  select(ac, port)

names(trial) = c("ac", "Community.Name")
trial<-trial %>%
  left_join(svi)

trial$Community.Name<-as.factor(trial$Community.Name)

trial<-trial[-c(16, 29, 43,52,59,61,64,80,93,96),]

#trial<-trial %>%
  #filter(!row_number() %in% c(16, 29, 43,52,59,61,64,80,93,96))

trial$ID=responses$X
domainsmod$ID=responses$X

sviscores <- trial %>%
  select(ID, Commercial.Engagement, Commercial.Reliance, Commercial.Engagement.Categorical.Ranking, Commercial.Reliance.Categorical.Ranking)

domainsmod<-domainsmod %>%
  left_join(sviscores)

domainsmod <- domainsmod %>%
  replace_na(list(Commercial.Engagement.Categorical.Ranking = 0, Commercial.Reliance.Categorical.Ranking = 0))

domainsmod$Commercial.Engagement.Categorical.Ranking<-as.numeric(domainsmod$Commercial.Engagement.Categorical.Ranking)
domainsmod$Commercial.Reliance.Categorical.Ranking<-as.numeric(domainsmod$Commercial.Reliance.Categorical.Ranking)

domainsmod <- domainsmod %>%
  mutate(dependence = (Commercial.Reliance.Categorical.Ranking + Commercial.Engagement.Categorical.Ranking))

domainsmod$dependence<-as.factor(domainsmod$dependence)
  
```

```{r}
ggplot(trial, aes(x = Commercial.Engagement, y = ac))+
  geom_point()

#trial$flexibility = domainsmod$flexibility
```
#Looking at all fisheries is kind of a mess, but everyone actually fits into one of these 4 categories:  salmon, dc, both, or neither
```{r}
domainsmod$crab = responses$dungeness_crab
domainsmod$salmon = responses$salmon

domainsmod <- domainsmod %>%
  mutate(fish_cat = case_when(
    salmon == 1 & crab != 1 ~ "salmon",
    salmon != 1 & crab == 1 ~ "crab",
    salmon == 1 & crab == 1 ~ "both",
    salmon != 1 & crab != 1 ~"neither")) 
```
#visualizing those fishery categories and the pca
```{r}
ggplot(domainsmod, aes(x = portlat, y = ac, color = fish_cat))+
  geom_point() +
  geom_smooth()

groupsfish <- as.factor(domainsmod$fish_cat)

fishgroups<-fviz_pca_ind(ac_dom_pca2,
             col.ind = groupsfish,
             #palette = pnw_palette("Bay", 62, type = "continuous"),
             addEllipses = TRUE, # Concentration ellipses
             ellipse.type = "confidence",
             legend.title = "Groups"
             )

fishgroups
```
#bar graph of each of those fishery categories and the average score in each domain of ac.
```{r}
fishsum<-domainsmod %>%
  group_by(fish_cat) %>%
  summarise(assets = mean(assets),
            #assetssd = sd(assets),
            flexibility = mean(flexibility),
            #flexibilitysd = sd(flexibility),
            organization = mean(organization),
            #organizationsd = sd(organization),
            learning = mean(learning),
            #learningsd = sd(learning),
            sociocog = mean(sociocog),
            #sociocogsd = sd(sociocog),
            agency = mean(agency))
            #agencysd = sd(agency))

fishsumlong<-melt(fishsum, id = "fish_cat")


fishbar<-ggplot(fishsumlong, aes(x=variable, y=value, fill = fish_cat)) +
  geom_bar(position = "dodge", stat = "identity")+
  scale_fill_manual(values = pal5)

fishbar
```

#clusters individuals based on their scores across the domains of ac. 
```{r}
NbClust(domainsmod[,1:6], distance = "euclidean", min.nc=2, max.nc=12, 
            method = "complete", index = "alllong")

#do the clustering
#heirachical with ward method
#computational costs = 4 mint milanos
d<-dist(domainsmod[,1:6], method = "euclidean")
con_cluster<-hclust(d, method = "ward")
groups<-cutree(con_cluster, k=3)
domainsmod<-cbind(domainsmod, groups)
domainsmod$groups<-as.factor(domainsmod$groups)

table(domainsmod$groups)

plot(con_cluster)
```
#Compares how the group membership varies by a few factors using chi squared tests.
```{r}
cc<-table(domainsmod$total_areas, domainsmod$groups)
chisq.test(cc)

dd<-table(domainsmod$groups, domainsmod$region)
chisq.test(dd)

ee<-table(domainsmod$groups, domainsmod$Commercial.Reliance.Categorical.Ranking)
chisq.test(ee)

ff<-table(domainsmod$groups, domainsmod$numfish)
chisq.test(ff)
ff

library("graphics")
mosaicplot(ff, shade = TRUE, las=2,
           main = "num fisheries")
```

#needs attention
```{r}
ggplot(domainsmod, aes(x = risk, y = ac, color = groups)) +
  geom_point()

dolong<-domainsmod %>%
  select(assets, flexibility, organization, learning, sociocog, agency, groups) %>%
  melt(id = "groups")

ggplot(dolong, aes())

ggplot(dolong, aes(x=variable, y=value, fill = groups)) +
  geom_bar(position = "dodge", stat = "identity")+
  scale_fill_manual(values = pal3)

groupstats<-domainsmod %>%
  group_by(groups) %>%
  summarise(assets = mean(assets),
            #assetssd = sd(assets),
            flexibility = mean(flexibility),
            #flexibilitysd = sd(flexibility),
            organization = mean(organization),
            #organizationsd = sd(organization),
            learning = mean(learning),
            #learningsd = sd(learning),
            sociocog = mean(sociocog),
            #sociocogsd = sd(sociocog),
            agency = mean(agency),
            ac = mean(ac))
            #agencysd = sd(agency))

write.csv(groupstats, paste0("../figures/ac/groupscluster.csv"))
```

Does the adaptive capacity vary among the clusters?
```{r}
#shit ton of anovas
#first overall ac
clus_aov<-aov(ac ~ groups, data = domainsmod)
summary(clus_aov)

#asumptions
library(car)
leveneTest(ac ~ groups, data = domainsmod)

clusresiduals<-residuals(object = clus_aov)
shapiro.test(x=clusresiduals)

#across domains
#assets
cass_aov<-aov(assets ~ groups, data = domainsmod)
summary(cass_aov)

#asumptions
leveneTest(assets ~ groups, data = domainsmod)

caov.residuals<-residuals(object = cass_aov)
shapiro.test(x=caov.residuals)

#assets meets levene test but fails shapiro, use kruskal wallis
kruskal.test(assets ~ groups, data = domainsmod)

#flexibilty
cflex_aov<-aov(flexibility ~ groups, data = domainsmod)
summary(cflex_aov)

#asumptions
leveneTest(flexibility ~ groups, data = domainsmod)

faov.residuals<-residuals(object = cflex_aov)
shapiro.test(x=faov.residuals)

kruskal.test(flexibility ~ groups, data = domainsmod)

#organization
corg_aov<-aov(organization ~ groups, data = domainsmod)
summary(corg_aov)

kruskal.test(organization ~ groups, data = domainsmod)

#asumptions
leveneTest(sociocog ~ region, data = domainsmod)

aov.residuals<-residuals(object = soc_aov)
shapiro.test(x=aov.residuals)

#learning
clearn_aov<-aov(learning ~ groups, data = domainsmod)
summary(clearn_aov)

kruskal.test(learning ~ groups, data = domainsmod)

#asumptions
leveneTest(sociocog ~ region, data = domainsmod)

aov.residuals<-residuals(object = soc_aov)
shapiro.test(x=aov.residuals)

#sociocog
csoc_aov<-aov(sociocog ~ groups, data = domainsmod)
summary(csoc_aov)

kruskal.test(sociocog ~ groups, data = domainsmod)

TukeyHSD(csoc_aov)

#agency
cagency_aov<-aov(agency ~ groups, data = domainsmod)
summary(cagency_aov)

kruskal.test(agency ~ groups, data = domainsmod)

TukeyHSD(cagency_aov)

#asumptions
leveneTest(sociocog ~ region, data = domainsmod)

aov.residuals<-residuals(object = soc_aov)
shapiro.test(x=aov.residuals)

#sociocog meets levene test but fails shapiro, use kruskal wallis
kruskal.test(sociocog ~ region, data = domainsmod)

#agency assumptions
leveneTest(agency ~ region, data = domainsmod)
aov.residuals<-residuals(object = agency_aov)
shapiro.test(x=aov.residuals)

#agency also fails run kruskal
kruskal.test(agency ~ region, data = domainsmod)
```
```

```{r}
domainsmod$ID = indv_risk$ID

domainsmod <- domainsmod %>%
  left_join(indv_risk)

ggplot(domainsmod, aes(x = ac, y = avg_risk, color = fish_cat)) +
  geom_point()

jjj<-table(domainsmod$groups, domainsmod$Commercial.Reliance.Categorical.Ranking)
chisq.test(jjj)
```