---
title: "Wellbeing"
author: "Laura"
date: "12/14/2020"
output: word_document
---


```{r setup, echo = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message = FALSE, error=FALSE)
```

```{r, packages}
library(tidyverse)
library(lubridate)
library(knitr)
library(grDevices)
library(extrafont)
library(ggrepel)
library(ggthemes)
library(gridExtra)
library(likert)
library(PNWColors)
# windowsFonts(Times=windowsFont("Calibri"))
theme_sleek <- function(base_size = 12, base_family = "Calibri") {
  half_line <- base_size/2
  theme_light(base_size = 12, base_family = "Calibri") +
    theme(
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      axis.ticks.length = unit(half_line / 2.2, "pt"),
      strip.background = element_rect(fill = NA, colour = NA),
      strip.text.x = element_text(colour = "black"),
      strip.text.y = element_text(colour = "black"),
      # axis.text = element_text(colour = "grey30"),
      #axis.title = element_text(colour = "grey30"),
      #legend.title = element_text(colour = "grey30"),#, size = rel(0.9)
      legend.title = element_blank(),
      panel.border = element_rect(fill = NA),#, colour = "grey70", size = 1),
      legend.key.size = unit(0.9, "lines"),
      #legend.text = element_text(size = rel(0.7)),#, colour = "grey30"),
      legend.key = element_rect(colour = NA, fill = NA),
      legend.background = element_rect(colour = NA, fill = NA)#,
      #plot.title = element_text(colour = "grey30"),#, size = rel(1),
      #plot.subtitle = element_text(colour = "grey30")#, size = rel(.85),
     
      
    )
}
theme_set(theme_sleek())
options(scipen = 999) #turns off scientific notation
# Depends on dplyr
tickr <- function(
  data, # dataframe
  var, # column of interest
  to # break point definition 
){
  
  VAR <- enquo(var) # makes VAR a dynamic variable
  
  data %>% 
    distinct(!!VAR) %>%
    ungroup(!!VAR) %>% 
    mutate(labels = ifelse(!!VAR %in% seq(to * round(min(!!VAR) / to), max(!!VAR), to),
                           !!VAR, "")) %>%
    select(breaks = UQ(VAR), labels)
}
```

```{r load data}
old_responses <- read.csv("../data/PNW_decoded_responses_wstate10.Feb.2021.csv")
```
##wellbeing indicators
##Below is a list of issues that may affect fishing success, your wellbeing, or the wellbeing of your community. For the following questions, please indicate your level of concern. Theme indicated in each section

```{r question 33}
#issues in the marine environement
responses$warming_concern = factor(responses$warming_concern, levels = c("very", "somewhat", "none"))
responses$oa_concern = factor(responses$oa_concern, levels = c("very", "somewhat", "none"))
responses$storm_concern = factor(responses$storm_concern, levels = c("very", "somewhat", "none"))
responses$sea_level_concern = factor(responses$sea_level_concern, levels = c("very", "somewhat", "none"))
responses$weather_concern = factor(responses$weather_concern, levels = c("very", "somewhat", "none"))
responses$h20_quality_concern = factor(responses$h20_quality_concern, levels = c("very", "somewhat", "none"))
responses$algal_concern = factor(responses$algal_concern, levels = c("very", "somewhat", "none"))
responses$habitat_concern = factor(responses$habitat_concern, levels = c("very", "somewhat", "none"))
         
warming <- ggplot(data = responses, aes(x=warming_concern))+
  geom_bar(stat = "count", fill = c("light blue", "navy blue", "purple"))+
  ylim(0, 90)+
  theme(axis.title.x = element_blank(), axis.title.y = element_blank(), axis.text.x = element_text(angle = 45),
        plot.title = element_text(size = 10))+ 
  ggtitle("ocean warming")

oa <- ggplot(data = responses, aes(x=oa_concern))+
  geom_bar(stat = "count", fill = c("light blue", "navy blue", "purple"))+
   ylim(0, 90)+
  theme(axis.title.x = element_blank(),axis.title.y = element_blank(), axis.text.x = element_text(angle = 45),
        plot.title = element_text(size = 10))+
  ggtitle("OA")

storm <- ggplot(data = responses, aes(x=storm_concern))+
  geom_bar(stat = "count", fill = c("light blue", "navy blue", "purple"))+
   ylim(0, 90)+ 
  theme(axis.title.x = element_blank(), axis.title.y = element_blank(), axis.text.x = element_text(angle = 45),
        plot.title = element_text(size = 10))+
  ggtitle("storms")

slr <- ggplot(data = responses, aes(x=sea_level_concern))+
  geom_bar(stat = "count", fill = c("light blue", "navy blue", "purple"))+
   ylim(0, 90)+
  theme(axis.title.x = element_blank(), axis.title.y = element_blank(), axis.text.x = element_text(angle = 45),
        plot.title = element_text(size = 10))+
  ggtitle("sea level rise")

weather <- ggplot(data = responses, aes(x=weather_concern))+
  geom_bar(stat = "count", fill = c("light blue", "navy blue", "purple"))+
   ylim(0, 90)+
  theme(axis.title.x = element_blank(), axis.title.y = element_blank(), axis.text.x = element_text(angle = 45),
        plot.title = element_text(size = 10))+
  ggtitle("Changing weather")

h20quality <- ggplot(data = responses, aes(x=h20_quality_concern))+
  geom_bar(stat = "count", fill = c("light blue", "navy blue", "purple"))+
   ylim(0, 90)+
  theme(axis.title.x = element_blank(), axis.title.y = element_blank(), axis.text.x = element_text(angle = 45),
        plot.title = element_text(size = 10))+
  ggtitle("water quality")

habs <- ggplot(data = responses, aes(x=algal_concern))+
  geom_bar(stat = "count", fill = c("light blue", "navy blue", "purple"))+
   ylim(0, 90)+
  theme(axis.title.x = element_blank(), axis.title.y = element_blank(), axis.text.x = element_text(angle = 45),
        plot.title = element_text(size = 10))+
  ggtitle("HABs")

habitat <- ggplot(data = responses, aes(x=habitat_concern))+
  geom_bar(stat = "count", fill = c("light blue", "navy blue", "purple"))+
   ylim(0, 90)+
  theme(axis.title.x = element_blank(), axis.title.y = element_blank(), axis.text.x = element_text(angle = 45),
        plot.title = element_text(size = 10))+
  ggtitle("habitat loss")

environment_concerns_plot<-grid.arrange(warming, oa, storm, slr, weather, h20quality, habs, habitat, nrow = 2, top = "Level of concern for issues in the marine environment", bottom = "concern level", left = "count")

ggsave(plot = environment_concerns_plot, file = paste0("../figures/wellbeing/environment_concerns.png"))

#visualizing same question with likert package

q33<-responses %>%
  select(warming_concern, oa_concern, storm_concern, sea_level_concern, weather_concern, h20_quality_concern, algal_concern, habitat_concern)

#add statements as column headers
q33_statements<-c(
  "ocean warming",
  "ocean acidification",
  "increases in storms",
  "sea level rise",
  "changing weather patterns",
  "ocean water quality",
  "harmful algal blooms",
  "habitat loss")
colnames(q33)<-q33_statements

#setting the levels
#levels need to be set since each possible answer does not appear in each column
mylevels <- c("very", "somewhat", "none")
for(i in seq_along(q33)) {
  q33[,i] <- factor(q33[,i], levels=mylevels)
}

ocean_concerns<-likert(q33)
summary(ocean_concerns)
ocean_concerns_likert<-plot(ocean_concerns, type = "heat")

ggsave(plot = ocean_concerns_likert, file = paste0("../figures/wellbeing/ocean_concerns.png"))
```

##issues that may affect fishing
```{r question 34}
#q34
responses$pop_size_concern = factor(responses$pop_size_concern, levels = c("very", "somewhat", "none"))
responses$bycatch_concern = factor(responses$bycatch_concern, levels = c("very", "somewhat", "none"))
responses$value_concern = factor(responses$value_concern, levels = c("very", "somewhat", "none"))
responses$costs_concern = factor(responses$costs_concern, levels = c("very", "somewhat", "none"))
responses$sa_concern = factor(responses$sa_concern, levels = c("very", "somewhat", "none"))
responses$distance_concern = factor(responses$distance_concern, levels = c("very", "somewhat", "none"))
responses$reg_concern = factor(responses$reg_concern, levels = c("very", "somewhat", "none"))
         
pop_size <- ggplot(data = responses, aes(x=pop_size_concern))+
  geom_bar(stat = "count", fill = c("light blue", "navy blue", "purple"))+
  ylim(0, 120)+
  theme(axis.title.x = element_blank(), axis.title.y = element_blank(), axis.text.x = element_text(angle = 45),
        plot.title = element_text(size = 10))+ 
  ggtitle("Fish populations")

bycatch <- ggplot(data = responses, aes(x=bycatch_concern))+
  geom_bar(stat = "count", fill = c("light blue", "navy blue", "purple"))+
   ylim(0, 120)+
  theme(axis.title.x = element_blank(),axis.title.y = element_blank(), axis.text.x = element_text(angle = 45),
        plot.title = element_text(size = 10))+
  ggtitle("Bycatch")

value <- ggplot(data = responses, aes(x=value_concern))+
  geom_bar(stat = "count", fill = c("light blue", "navy blue", "purple"))+
   ylim(0, 120)+ 
  theme(axis.title.x = element_blank(), axis.title.y = element_blank(), axis.text.x = element_text(angle = 45),
        plot.title = element_text(size = 10))+
  ggtitle("Landed value")

costs <- ggplot(data = responses, aes(x=costs_concern))+
  geom_bar(stat = "count", fill = c("light blue", "navy blue", "purple"))+
   ylim(0, 120)+
  theme(axis.title.x = element_blank(), axis.title.y = element_blank(), axis.text.x = element_text(angle = 45),
        plot.title = element_text(size = 10))+
  ggtitle("Operational costs")

stockass <- ggplot(data = responses, aes(x=sa_concern))+
  geom_bar(stat = "count", fill = c("light blue", "navy blue", "purple"))+
   ylim(0, 120)+
  theme(axis.title.x = element_blank(), axis.title.y = element_blank(), axis.text.x = element_text(angle = 45),
        plot.title = element_text(size = 10))+
  ggtitle("Stock assessment")

distance <- ggplot(data = responses, aes(x=distance_concern))+
  geom_bar(stat = "count", fill = c("light blue", "navy blue", "purple"))+
   ylim(0, 120)+
  theme(axis.title.x = element_blank(), axis.title.y = element_blank(), axis.text.x = element_text(angle = 45),
        plot.title = element_text(size = 10))+
  ggtitle("Travel time")

regs <- ggplot(data = responses, aes(x=reg_concern))+
  geom_bar(stat = "count", fill = c("light blue", "navy blue", "purple"))+
   ylim(0, 120)+
  theme(axis.title.x = element_blank(), axis.title.y = element_blank(), axis.text.x = element_text(angle = 45),
        plot.title = element_text(size = 10))+
  ggtitle("Regulations")

fishing_concerns_plot<-grid.arrange(pop_size, bycatch, value, costs, stockass, distance, regs, nrow = 2, top = "Level of concern for issues related to fishing", bottom = "concern level", left = "count")

ggsave(plot = fishing_concerns_plot, file = paste0("../figures/wellbeing/fishing_concerns.png"))
```

##labor force and health
```{r questions 35 and 36}
#q35 and q36

responses$graying_fleet_concern = factor(responses$graying_fleet_concern, levels = c("very", "somewhat", "none"))
responses$fish_comm_concern = factor(responses$fish_comm_concern, levels = c("very", "somewhat", "none"))
responses$resid_concern = factor(responses$resid_concern, levels = c("very", "somewhat", "none"))
responses$infrastructure_concern = factor(responses$infrastructure_concern, levels = c("very", "somewhat", "none"))
responses$phsy_health_concern = factor(responses$phsy_health_concern, levels = c("very", "somewhat", "none"))
responses$mental_health_quality_concern = factor(responses$mental_health_concern, levels = c("very", "somewhat", "none"))
responses$safe_at_sea_concern = factor(responses$safe_at_sea_concern, levels = c("very", "somewhat", "none"))
responses$family_concern = factor(responses$family_concern, levels = c("very", "somewhat", "none"))
         
graying_fleet <- ggplot(data = responses, aes(x=graying_fleet_concern))+
  geom_bar(stat = "count", fill = c("light blue", "navy blue", "purple"))+
  ylim(0, 90)+
  theme(axis.title.x = element_blank(), axis.title.y = element_blank(), axis.text.x = element_text(angle = 45),
        plot.title = element_text(size = 10))+ 
  ggtitle("Graying of the fleet")

fish_comm <- ggplot(data = responses, aes(x=fish_comm_concern))+
  geom_bar(stat = "count", fill = c("light blue", "navy blue", "purple"))+
   ylim(0, 90)+
  theme(axis.title.x = element_blank(),axis.title.y = element_blank(), axis.text.x = element_text(angle = 45),
        plot.title = element_text(size = 10))+
  ggtitle("Fishing community")

resid_comm <- ggplot(data = responses, aes(x=resid_concern))+
  geom_bar(stat = "count", fill = c("light blue", "navy blue", "purple"))+
   ylim(0, 90)+ 
  theme(axis.title.x = element_blank(), axis.title.y = element_blank(), axis.text.x = element_text(angle = 45),
        plot.title = element_text(size = 10))+
  ggtitle("Residential community")

infrastructure <- ggplot(data = responses, aes(x=infrastructure_concern))+
  geom_bar(stat = "count", fill = c("light blue", "navy blue", "purple"))+
   ylim(0, 90)+
  theme(axis.title.x = element_blank(), axis.title.y = element_blank(), axis.text.x = element_text(angle = 45),
        plot.title = element_text(size = 10))+
  ggtitle("Infrastructure")

phsy_health <- ggplot(data = responses, aes(x=phsy_health_concern))+
  geom_bar(stat = "count", fill = c("light blue", "navy blue", "purple"))+
   ylim(0, 90)+
  theme(axis.title.x = element_blank(), axis.title.y = element_blank(), axis.text.x = element_text(angle = 45),
        plot.title = element_text(size = 10))+
  ggtitle("Physical health")

mental_health <- ggplot(data = responses, aes(x=mental_health_concern))+
  geom_bar(stat = "count", fill = c("light blue", "navy blue", "purple"))+
   ylim(0, 110)+
  theme(axis.title.x = element_blank(), axis.title.y = element_blank(), axis.text.x = element_text(angle = 45),
        plot.title = element_text(size = 10))+
  ggtitle("Mental health")

sas <- ggplot(data = responses, aes(x=safe_at_sea_concern))+
  geom_bar(stat = "count", fill = c("light blue", "navy blue", "purple"))+
   ylim(0, 90)+
  theme(axis.title.x = element_blank(), axis.title.y = element_blank(), axis.text.x = element_text(angle = 45),
        plot.title = element_text(size = 10))+
  ggtitle("Safety at sea")

family_rel <- ggplot(data = responses, aes(x=family_concern))+
  geom_bar(stat = "count", fill = c("light blue", "navy blue", "purple"))+
   ylim(0, 90)+
  theme(axis.title.x = element_blank(), axis.title.y = element_blank(), axis.text.x = element_text(angle = 45),
        plot.title = element_text(size = 10))+
  ggtitle("Family relationships")

community_concerns_plot<-grid.arrange(graying_fleet, fish_comm, resid_comm, infrastructure, phsy_health, mental_health, sas, family_rel, nrow = 2, top = "Level of concern for labor, health, and relationships", bottom = "concern level", left = "count")

ggsave(plot = community_concerns_plot, file = paste0("../figures/wellbeing/community_concerns.png"))
```

##comparing all concerns together
```{r}
all_concerns<-responses %>%
  select(warming_concern, oa_concern, storm_concern, sea_level_concern, weather_concern, h20_quality_concern, algal_concern, habitat_concern, pop_size_concern, bycatch_concern, value_concern, costs_concern, sa_concern, distance_concern, reg_concern, graying_fleet_concern, fish_comm_concern, resid_concern, infrastructure_concern, phsy_health_concern, mental_health_concern, safe_at_sea_concern, family_concern)

#add statements as column headers
all_concerns_statements<-c(
  "ocean warming",
  "ocean acidification",
  "increases in storms",
  "sea level rise",
  "changing weather patterns",
  "ocean water quality",
  "harmful algal blooms",
  "habitat loss",
  "fish populations",
  "bycatch",
  "landed value",
  "operational costs",
  "stock assessments",
  "increased travel distance",
  "regulations",
  "graying of the fleet",
  "fishing community cohesion",
  "residential community cohesion",
  "infrastructure",
  "physical health",
  "mental health",
  "safety at sea",
  "family")
colnames(all_concerns)<-all_concerns_statements

#setting the levels
#levels need to be set since each possible answer does not appear in each column
mylevels <- c("very", "somewhat", "none")
for(i in seq_along(all_concerns)) {
  all_concerns[,i] <- factor(all_concerns[,i], levels=mylevels)
}

total_concerns<-likert(all_concerns)

total_concerns_likert<-plot(total_concerns, type = "heat")+
  ggtitle("Environmental, social, and operational concerns")

ggsave(plot = total_concerns_likert, file = paste0("../figures/wellbeing/all_concerns.png"))

```

##likert questions about climate and future of fishing
```{r question 37}
#q37 plotting using likert package

q37<-responses %>%
  select(believe, harm_me, harm_future, leave_fishing, risk, no_point, no_fish)

#removes _ so legend looks nicer in figure
q37<-q37 %>%
  mutate(across(everything(), ~ case_when(
    (. == "strongly_disagree" ~ "strongly disagree"),
    (. == "somewhat_disagree" ~ "somewhat disagree"),
    (. == "neutral" ~ "neutral"),
    (. == "somewhat_agree" ~ "somewhat agree"),
    (. == "strongly_agree" ~ "strongly agree")
  )))

#add statements as column headers
q37_statements<-c(
  "I believe climate change is occurring",
  "Climate change will harm me personally",
  "Climate change will harm future generations",
  "If I had a choice I would leave fishing",
  "It is a big risk to move into a new fishery",
  "There is no point in preparing for climate change since we don't know exactly what is going to happen",
  "There will not be enough fish to continue to operate in my main fishery in 20 years")
colnames(q37)<-q37_statements


#setting the levels
#levels need to be set since each possible answer does not appear in each column
mylevels <- c('strongly disagree', 'somewhat disagree', 'neutral', 'somewhat agree', 'strongly agree')
for(i in seq_along(q37)) {
  q37[,i] <- factor(q37[,i], levels=mylevels)
}

future_thoughts<-likert(q37)
summary(future_thoughts)
climate_perspectives<-plot(future_thoughts,   colors=rev(pnw_palette("Sailboat", 5))) + theme(axis.text = element_text(size = 12))

climate_perspectives

ggsave(plot = climate_perspectives, height = 7, width = 10, file = paste0("../figures/wellbeing/climate_perspectives.png"))

#by state
#future_thoughts_states<-likert(q37, grouping = responses$state_code)
#climate_perspectives_states_plot<-plot(future_thoughts_states, colors=c('orange','darkorange', "gray", 'darkblue','blue'))

#ggsave(plot = climate_perspectives_states_plot, height = 12, width = 10, file = paste0("../figures/wellbeing/climate_perspectives_state.png"))

```

##likert questions about conflict
```{r question 38}
#q38 plotting using likert package

q38<-responses %>%
  select(recreational, other_fishery, internal, aquaculture, hatcheries, tourism, development, spatial)

#add statements as column headers
q38_statements<-c(
  "Recreational fisheries",
  "Other commerical fisheries",
  "There is internal conflict within my own commercial fishery",
  "Aquaculture",
  "Hatcheries",
  "Tourism",
  "Coastal development",
  "Competing ocean uses or offshore energy development")
colnames(q38)<-q38_statements

#setting the levels
#levels need to be set since each possible answer does not appear in each column
mylevels <- c('strongly_disagree', 'somewhat_disagree', 'neutral', 'somewhat_agree', 'strongly_agree')
for(i in seq_along(q38)) {
  q38[,i] <- factor(q38[,i], levels=mylevels)
}

conflict<-likert(q38)
summary(conflict)
conflict_title<-"My fishing is negatively affected by:"
conflict_plot<-plot(conflict, 
              colors=c('orange','darkorange', "gray", 'darkblue','blue')) +       ggtitle(conflict_title)

ggsave(plot = conflict_plot, height = 7, width = 10, file = paste0("../figures/wellbeing/conflict_plot.png"))

#by state
conflict_states<-likert(q38, grouping = responses$state_code)
conflict_states_plot<-plot(conflict_states, colors=c('orange','darkorange', "gray", 'darkblue','blue')) + ggtitle(conflict_title)

ggsave(plot = conflict_states_plot, height = 12, width = 10, file = paste0("../figures/wellbeing/conflict_state.png"))

```
##likert questions about health and wellbeing
```{r question 39}
#q39 plotting using likert package
#conditions indicators in wellbeing framework

q39<-old_responses %>%
  select(fish_neg_wellbeing, fish_neg_health, fish_neg_mental, fish_increase_stress, env_neg_welbeing, env_neg_health, env_neg_mental, env_increase_stress, env_neg_safety)

#add statements as column headers
q39_statements<-c(
  "Changes in fisheries have negatively impacted my overall wellbeing",
  "Changes in fisheries have negatively impacted my physical health",
  "Changes in fisheries have negatively impacted my mental health",
  "Changes in fisheries have raised my stress levels",
  "Changes in the environment have negatively impacted my overall wellbeing",
  "Changes in the environment have negatively impacted my physical health",
  "Changes in the environment have negatively impacted my mental health",
  "Changes in the environment have raised my stress levels",
  "Changes in the environment have negatively affected my safety while fishing")
colnames(q39)<-q39_statements

#setting the levels
#levels need to be set since each possible answer does not appear in each column
mylevels <- c('strongly_disagree', 'somewhat_disagree', 'neutral', 'somewhat_agree', 'strongly_agree')
for(i in seq_along(q39)) {
  q39[,i] <- factor(q39[,i], levels=mylevels)
}

health_wellbeing<-likert(q39)
health_wellbeing_title<-"Level of agreement with the following statements:"
health_wellbeing_plot<-plot(health_wellbeing, 
              colors=rev(pnw_palette("Sailboat", 5))) +       
  ggtitle(health_wellbeing_title)

ggsave(plot = health_wellbeing_plot, height=7, width=10, file = paste0("../figures/wellbeing/health_wellbeing_plot.png"))

#by state
health_states<-likert(q39, grouping = responses$state_code)
health_states_plot<-plot(health_states, colors=c('orange','darkorange', "gray", 'darkblue','blue')) + ggtitle(health_wellbeing_title)

ggsave(plot = health_states_plot, height = 12, width = 10, file = paste0("../figures/wellbeing/health_state.png"))
```
##Below is a list of statements that may indicate the degree to which community and individual well-being is sensitive to changes in the health of fisheries and the environment. Please indicate your level of agreement with each statement.
```{r question 40}
#q40 plotting using likert package
#connections indicators in wellbeing framework

q40<-responses %>%
  select(enc_child, identity, comm_support, generation, connect_env, connect_comm)

#add statements as column headers
q40_statements<-c(
  "I would encourage my children to be fishermen",
  "Fishing in important to my identity",
  "Fishermen are supported in my community",
  "I am passing down fishing knowledge to the next generation",
  "I feel a connection to my environment",
  "I feel a connection to my community")
colnames(q40)<-q40_statements

#setting the levels
#levels need to be set since each possible answer does not appear in each column
mylevels <- c('strongly_disagree', 'somewhat_disagree', 'neutral', 'somewhat_agree', 'strongly_agree')
for(i in seq_along(q40)) {
  q40[,i] <- factor(q40[,i], levels=mylevels)
}

connections<-likert(q40)
connections_title<-"Level of agreement with the following statements:"
connections_plot<-plot(connections, 
              colors=c('orange','darkorange', "gray", 'darkblue','blue')) +       
  ggtitle(connections_title)

ggsave(plot = connections_plot, height=7, width=10, file = paste0("../figures/wellbeing/connections_plot.png"))
```

```{r question 41}
#q41 plotting using likert package
#capabilities indicators in wellbeing framework

q41<-responses %>%
  select(money, plan, job_sat, effective_mgmt, data, find_crew, voice, climate_in_mgmt)

#add statements as column headers
q41_statements<-c(
  "I make enough money to support my family",
  "I am able to plan two years out in the future",
  "I am satisfied with my job",
  "I think the fisheries I participate in are managed effectively",
  "I have access to the data and information I need for successful fishing",
  "I can find qualified crew with the skills they need to do a good job",
  "I have a voice in fisheries management",
  "Climate change should be considered in fisheries management")
colnames(q41)<-q41_statements

#setting the levels
#levels need to be set since each possible answer does not appear in each column
mylevels <- c('strongly_disagree', 'somewhat_disagree', 'neutral', 'somewhat_agree', 'strongly_agree')
for(i in seq_along(q41)) {
  q41[,i] <- factor(q41[,i], levels=mylevels)
}

capabilities<-likert(q41)
capabilities_title<-"Level of agreement with the following statements:"
capabilities_plot<-plot(capabilities, 
              colors=c('orange','darkorange', "gray", 'darkblue','blue')) +       
  ggtitle(capabilities_title)

ggsave(plot = capabilities_plot, height=7, width=10, file = paste0("../figures/wellbeing/capabilities_plot.png"))
```

```{r question 42}
#q42 plotting using likert package
#cross-cutting indicators in wellbeing framework

q42<-responses %>%
  select(equitable, opp_outside, opp_crew)

#add statements as column headers
q42_statements<-c(
  "The fisheries I participate in are managed in an equitable way",
  "There are opportunities for people who are not currently fishing to enter into west coast fisheries",
  "There are opportunities for deckhands and other lower level crew to advance")
colnames(q42)<-q42_statements

#setting the levels
#levels need to be set since each possible answer does not appear in each column
mylevels <- c('strongly_disagree', 'somewhat_disagree', 'neutral', 'somewhat_agree', 'strongly_agree')
for(i in seq_along(q42)) {
  q42[,i] <- factor(q42[,i], levels=mylevels)
}

crosscutting<-likert(q42)
crosscutting_title<-"Level of agreement with the following statements:"
crosscutting_plot<-plot(crosscutting, 
              colors=c('orange','darkorange', "gray", 'darkblue','blue')) +       
  ggtitle(crosscutting_title)

ggsave(plot = crosscutting_plot, height=7, width=10, file = paste0("../figures/wellbeing/crosscutting_plot.png"))
```
##With regard to the future security of yourself, your residential community, or your fishery, please indicate the extent to which you agree or disagree with the following statements:
```{r question 43}
#q43 plotting using likert package
#adapative capacity indicators

q43<-old_responses %>%
  select(new_fishery, nat_resource_job, other_job, loan, distance, strong_comm, climate_refugees, adapt_mgmt, const_reg)

#add statements as column headers
q43_statements<-c(
  "I could easily move into a new fishery",
  "I could easily find work in another natural resource industry (aquaculture, forestry, etc.)",
  "I could easily get income not related to natural resource harvest, fishing or otherwise",
  "I could easily get a loan or some other form of financial support",
  "I am confident in my ability to travel further to fish if that is needed",
  "I believe my community has a strong and viable future ahead",
  "I am concerned that climate change may lead to people moving out of my community",
  "I think fisheries management can adapt and respond quickly to changing environmental conditions",
  "I feel constrained in my ability to adapt to changes because of regulations")
colnames(q43)<-q43_statements

#setting the levels
#levels need to be set since each possible answer does not appear in each column
mylevels <- c('strongly_disagree', 'somewhat_disagree', 'neutral', 'somewhat_agree', 'strongly_agree')
for(i in seq_along(q43)) {
  q43[,i] <- factor(q43[,i], levels=mylevels)
}

ac_likert<-likert(q43)
ac_likert_title<-"Level of agreement with the following statements:"
ac_likert_plot<-plot(ac_likert, 
              colors=rev(pnw_palette("Sailboat", 5))) +       
  ggtitle(ac_likert_title)

ggsave(plot = ac_likert_plot, height=7, width=10, file = paste0("../figures/wellbeing/ac_likert_plot.png"))

#by state
ac_states<-likert(q43, grouping = responses$state_code)
ac_states_plot<-plot(ac_states, colors=c('orange','darkorange', "gray", 'darkblue','blue')) + ggtitle(ac_likert_title)

ggsave(plot = ac_states_plot, height = 12, width = 10, file = paste0("../figures/wellbeing/ac_state.png"))
```

##select concerns for presentation
```{r}
some_concerns<-responses %>%
  select(warming_concern, sea_level_concern,  algal_concern, habitat_concern, pop_size_concern, bycatch_concern, costs_concern, reg_concern, safe_at_sea_concern)

#add statements as column headers
some_concerns_statements<-c(
  "ocean warming",
  "sea level rise",
  "harmful algal blooms",
  "habitat loss",
  "fish populations",
  "bycatch",
  "operational costs",
  "regulations",
  "safety at sea")
colnames(some_concerns)<-some_concerns_statements

#setting the levels
#levels need to be set since each possible answer does not appear in each column
mylevels <- c("very", "somewhat", "none")
for(i in seq_along(some_concerns)) {
  some_concerns[,i] <- factor(some_concerns[,i], levels=mylevels)
}

tnc_concerns<-likert(some_concerns)

tnc_concerns_likert<-plot(tnc_concerns, type = "heat")+
  ggtitle("Environmental and operational concerns")+
  theme(axis.text=element_text(size=14), plot.title = element_text(size = 16))
tnc_concerns_likert

ggsave(plot = tnc_concerns_likert, file = paste0("../figures/wellbeing/tnc_concerns.png"))

```

#Looking at select questions here for adaptive capcity paper
```{r question 43}
#q43 plotting using likert package
#adapative capacity indicators

ac<-old_responses %>%
  select(new_fishery, nat_resource_job, other_job)

#removes _ so legend looks nicer in figure
ac<-ac %>%
  mutate(across(everything(), ~ case_when(
    (. == "strongly_disagree" ~ "strongly disagree"),
    (. == "somewhat_disagree" ~ "somewhat disagree"),
    (. == "neutral" ~ "neutral"),
    (. == "somewhat_agree" ~ "somewhat agree"),
    (. == "strongly_agree" ~ "strongly agree")
  )))

#add statements as column headers
ac_statements<-c(
  "I could easily move into a new fishery",
  "I could easily find work in another natural resource industry (aquaculture, forestry, etc.)",
  "I could easily get income not related to natural resource harvest, fishing or otherwise")
colnames(ac)<-ac_statements

#setting the levels
#levels need to be set since each possible answer does not appear in each column
mylevels <- c('strongly disagree', 'somewhat disagree', 'neutral', 'somewhat agree', 'strongly agree')
for(i in seq_along(ac)) {
  ac[,i] <- factor(ac[,i], levels=mylevels)
}

ac_likert<-likert(ac)
ac_likert_title<-"Level of agreement with the following statements:"
ac_likert_plot<-plot(ac_likert, 
              colors=rev(pnw_palette("Sailboat", 5))) +       
  ggtitle(ac_likert_title)+ theme(axis.text = element_text(size = 12))

ggsave(plot = ac_likert_plot, height=4, width=8, file = paste0("../figures/ac/ac_paper_plot.png"))

#by region
old_responses$region = domainsmod$region
ac_region<-likert(ac, grouping = old_responses$region)
ac_region_plot<-plot(ac_region, colors=rev(pnw_palette("Sailboat", 5))) + ggtitle(ac_likert_title)+ theme(axis.text = element_text(size = 12))

ac_region_plot

ggsave(plot = ac_region_plot, height = 12, width = 10, file = paste0("../figures/ac/ac_state.png"))
```

#spliting out the travel question
```{r question 43}
#q43 plotting using likert package
#adapative capacity indicators

travel<-ac %>%
  select("I am confident in my ability to travel further to fish if that is needed")


#add statements as column headers
#travel_statements<-c(
  #"I am confident in my ability to travel further to fish if that is needed")
#colnames(ac)<-travel_statements

#setting the levels
#levels need to be set since each possible answer does not appear in each column
mylevels <- c('strongly disagree', 'somewhat disagree', 'neutral', 'somewhat agree', 'strongly agree')
for(i in seq_along(travel)) {
  travel[,i] <- factor(travel[,i], levels=mylevels)
}

travel_likert<-likert(travel)
travel_likert_title<-"Level of agreement with the following statements:"
travel_likert_plot<-plot(travel_likert, 
              colors=rev(pnw_palette("Sailboat", 5))) +       
  ggtitle(travel_likert_title)+ theme(axis.text = element_text(size = 12))

ggsave(plot = travel_likert_plot, height=7, width=10, file = paste0("../figures/ac/travel_plot.png"))

#by region
old_responses$region = domainsmod$region
travel_region<-likert(travel, grouping = old_responses$region)
travel_region_plot<-plot(travel_region, colors=rev(pnw_palette("Sailboat", 5))) + ggtitle(travel_likert_title)+ theme(axis.text = element_text(size = 12))

travel_region_plot

ggsave(plot = travel_region_plot, height = 5, width = 7.5, file = paste0("../figures/ac/ac_travel.png"))
```

